% -*- mode: LaTeX; -*-


\chapter*{People}

\section*{Authors}

The original \emph{Europa Universalis} game was written by Philippe Thibaud
for \emph{Azure Wish Edition} in 1993. Risto Marjomaa did a tremendous work of
rewriting and clarifying all the events.

This rewrite is based both on the original version and the rewrite of the
events and include many new things. Numerous comments from the EU mailing-list
were also taken into account when designing the map and balancing details of
the rules.

This text was written by the united minds of Pierre Borgnat, Bertrand Asseray,
Jean-Yves Moyen and Jean-Christophe Dubacq. Composition was made using
\TeX. Accompanying maps were created by J.-C. Dubacq. Illustrations used in
chapter headers are excerpts of famous masterpieces as follow:
% Basics
 \emph{The Astronomer}, Vermeer;
% The powers
\emph{Seaport}, Lorrain;
% Events (rules)
\emph{La liberté guidant le peuple}, Delacroix;
% Diplomacy
the American Constitution;
% Incomes
\emph{Harvesters}, Brueghel the Elder;
% Expenses
\emph{Construction of the Château de Versailles}, Meulen;
% Military
\emph{Night Watch}, Rembrandt;
% Peaces
\emph{The surrender of Breda}, Velázquez;
% Victories
\emph{Le sacre de Napoléon}, Joseph-Louis David;
% Specific rules
\emph{Le joueur}, de la Tour;
% Events (description)
\emph{Bodas de Caná} Murillo;
% Political events
\emph{The Creation of Adam}, Michelangelo;
% Appendix minors
manuscript of \emph{Beowulf};
% Appendix majors
\emph{World map}, Mercator;
% Scenarios
\emph{Columbus landing}, John Vanderlyn; 
% Playing the game
woodcut from an Italian chess treatise (1493);
% Strategy
\emph{The Cardsharps}, Carravagio;
% ???
picture of the Vatican Library.




\section*{Testers}
% in alphabetical order

Preliminary versions of the game were tested by:\par
Bertrand Asseray, Pierre Borgnat, Nicolas Bourgeois, Michael ``MKL'' Chagnon,
Silvère Corsange, Sylvain Domergue, Jean-Christophe Dubacq, Manuel Esteban,
Erwan ``R1'' Grasland-Mongrain, Maxime le Heiget, Bruno ``Boka'' Kauffmann,
Gwendal Kervern, Benoît ``Kloecky'' Kloeckner, Thomas ``Satori'' Milot,
Jean-Yves ``Jym'' Moyen, Rémi Moyen, Fabrice Randeau, Éric ``Troll''
Schaeffler, Aurélien Schoumaker, Benoît Trédez and Benjamin Wack.

% A first game of EU8 was held between 2002 and 2004. It was played by Pierre
% Borgnat (Spain), Bertrand Asseray (France), Jean-Yves ``Jym'' Moyen
% (Portugal-Sweden), Sylvain Domergue and Jean-Christophe Dubacq (Russia),
% Aurélien Schoumaker (Venice-Holland-Austria), Rémi Moyen (England), Benjamin
% Wack (Turkey) and Éric ``Troll'' Schaeffler (Poland-Prussia)

% A second game was held between 2005 and 2007 featuring Pierre Borgnat
% (Portugal/Poland-Prussia), Bertrand Asseray (Turkey), Jean-Yves ``Jym''
% Moyen (France), Jean-Christophe Dubacq (Spain), Aurélien Schoumaker
% (Venice-Holland-Austria), Benoît ``Kloecky'' Kloeckner (England), Rémi Moyen
% and Fabrice Randeau (Russia) and Nicolas Bourgeois (Sweden).

% A third game was held between 2006 and 2008 with Jean-Yves ``Jym'' Moyen
% (Portugal-Sweden), Nicolas Bourgeois (Spain), Erwan ``R1'' Grasland-Mongrain
% (Russia), Manuel Esteban (France), Silvère Corsange (England), Michael
% ``MKL'' Chagnon (Venice-Holland-Austria), Bruno ``Boka'' Kauffmann (Turkey)
% and Thomas ``Satori'' Milot (Poland-Prussia).

% A fourth game started in 2008 with Pierre Borgnat (Venice-Austria), Bertrand
% Asseray (England), Jean-Yves ``Jym'' Moyen (Poland-Prussia), Jean-Christophe
% Dubacq (Holland), Aurélien Schoumaker (Spain), Nicolas Bourgeois (France),
% Michael ``MKL'' Chagnon (Russia), Bruno ``Boka'' Kauffmann (Portugal-Sweden)
% and Thomas ``Satori'' Milot (Turkey).

% A fifth game started in 2010 with Jean-Yves ``Jym'' Moyen (Turkey), Nicolas
% Bourgeois (Venice-Austria), Michael ``MKL'' Chagnon (Spain), Thomas
% ``Satori'' Milot (France), Bruno ``Boka'' Kauffmann (England), Gwendal
% Kervern (Holland), Maxime le Heiget (Portugal-Sweden), Benoît Trédez
% (Poland-Prussia) and Silvère Corsange (Russia).


% Local Variables:
% fill-column: 78
% coding: utf-8-unix
% mode-require-final-newline: t
% mode: flyspell
% ispell-local-dictionary: "british"
% End:
