% -*- mode: LaTeX; -*-
\chapter{Political Events of Period IV}
%\section{Period IV}
\label{events:pIV}



\subsection*{Event Table of Period IV}

\begin{eventstable}[Period IV events table]
  \tabcolsep=4.5pt\centering%
  \begin{tabular}{|l|*{6}{c}|l|}
    \hline
    1\up{st}\textarrow& 1-3 & 4-5 & 6 & 7 & 8 & 9 & 10 \\ \hline
    1 & 1  &  1 & 13 & R4 & R19 & 7   & \textbullet~1--2: \\
    2 & 12 & 14 & 15 & R5 & 18  & 8   & +1 then\\
    3 & 17 & 15 & 9  & 6  & 17  & R9  & \nameref{events:pIII}\\
    4 & 18 & 16 & 10 & R7 & 16  & R17 & \textbullet~3--10:\\
    5 & 10 &  4 & R11& 8  & 14  & R18 & \nameref{events:pIII}\\
    6 & 3  &  2 & 12 & 9  & 1   & 19  & \\
    7 & 7  &  6 & 1  & 11 & R5  & R20 & \\
    8 & 22 & R4 & 2  & 12 & 21  & R4  & \\
    9 & 5  & R7 & 3  & 13 & R22 & 8   & \\ \hline
    10 & \multicolumn{7}{l|}{\nameref{events:pV}} \\ \hline
  \end{tabular}
\end{eventstable}

\eventssummary{%
  pIV:Bohemian Revolt|,%
  pIV:Augsburg Revocation|,%
  pIV:Olivares|,%
  pIV:Unity HRE|,%
  pIV:War Persia Turkey|,%
  pIV:Persian Safavids|O{pIII:Persian Safavids},%
  % Change to Morocco/Portuguese Revolt ?
  pIV:Portuguese Revolt|,%
  pIV:Morocco|,%
  pIV:Act Navigation|,%
  pIV:Union Scotland|,%
  pIV:English Civil War|,%
  pIV:English Restoration|,%
  pIV:London Stock Exchange|,%
  pIV:Amsterdam Stock Exchange|O{pIII:Amsterdam Stock Exchange},%
  pIV:Dutch Colonial Dynamism|E/E/E,%
  pIV:Liberum Veto|,%
  pIV:Great Elector|,%
} \eventssummary{%
  pIV:Oxenstierna|O{pIII:Oxenstierna},%
  pIV:Union Poland Sweden|O{pIII:Union Poland Sweden},%
  pIV:Torstensson|,%
  pIV:Swedish Nobles Unrest|,%
  pIV:La Rochelle|E/O{pIV:Morocco},%
  pIV:Richelieu|,%
  pIV:Fronde|,%
  pIV:Times of Troubles|,%
  pIV:Revolt Cossacks|,%
  pIV:Extension Moghol|E/E,%
  pIV:Wars India|E/E,%
  pIV:Revolt Singala|E/E,%
  pIV:China Colonial Attitude|O{pIII:China Colonial
    Attitude}/S{pIV:CCA:Vassalisation Korea},%
  pIV:Japan Colonial Attitude|,%
  pIV:Deluge|,%
  pIV:Koprulu|,%
  O|,%
  pIV:TYW|,%
  pIV:Polish Civil War|,%
}

\newpage\startevents



\event{pIV:Bohemian Revolt}{IV-1 (1)}{Bohemian Revolt}{1}{PBNew}

\history{1618-1621}[This event describes the War for \paysBoheme, whereas the
break out of a general German conflict (that historically followed this event)
is dealt with in \ref{pIV:TYW}.]

\phevnt
\aparag[The Winter King] The minor country \paysBoheme is created / separated
/ breaks alliance (depending on its previous status) from its current
allegiance (even a \GE), and allies itself with \paysPalatinat (which would
also breaks from an existing \GE). The first Major power in the list: \FRA
(except if \CATHCR), \POL (if Protestant), else \SUE (even if Catholic)
controls both those countries and have them placed in \EG on its diplomatic
track.
\aparag \paysBaviere and \hab declares war to these two countries. This is a
Religious Civil War (see \ref{chDiplo:Religious Civil War}) in the \HRE.
\bparag \AUS has a free \CB against \paysBoheme and must use it or lose {\bf
  2} \STAB.
% (JCD) probably, \paysBaviere should be controlled by \MAJHAB, not
% \SPA. Changing.
\bparag If \HAB declares war, \paysBaviere is placed in \EG of \HAB and is
controlled by \MAJHAB.
\bparag \SPA is allowed a limited intervention in the war as an ally of
\HAB. Other countries are constrained by usual rules.

\aparag[The Revolt of Bethlén]
\bparag A \REVOLT \faceplus is placed in a randomly chosen province belonging
to \HAB in \paysHongrie. It controls the city. The \REVOLT is controlled by
\RUS.
\bparag The military forces of the Revolt of Bethlén can use up to 1 \ARMY and
2 \LD of the Hungarian counters (and the \HAB can use at most one \ARMY and 2
\LD from Hungarian counters).
\aparag \TUR cannot declare war against \HAB at this turn.

\phadm
\aparag \AUSMin receives its usual forces and reinforcements.
\aparag \paysBaviere has 1 \ARMY \faceplus, 3 \LD (all Veterans), 1 \fortress
and is commanded by \leaderwithdata{Tilly} (lasting 4 turns).  It has 2
Multiple Campaigns. \paysBaviere has 2 \ARMY counters at its disposition
during the whole length of this event.
\bparag[Tilly's training] [BLP] Troops of \paysBaviere (not its allies)
stacked with \leaderTilly are always \TTER.
\aparag \paysBoheme has 1 \ARMY \faceplus, 1 \LD (Conscripts) and 1 \fortress.
\aparag \paysPalatinat has 1 \ARMY \faceplus (Veterans) leaded by
\leaderwithdata{Mansfeld} (lasting 3 turns). It has 1 Multiple Campaign.
\aparag The Revolt of Bethlén consists of one Hungarian \ARMY \faceplus
(Conscripts) and \leaderwithdata{Bethlen} (lasting 4 turns) placed in the
province of the \REVOLT .
\aparag None of \paysBaviere, \paysBoheme, \paysPalatinat and the Revolt of
Bethlén receive reinforcements on the first turn. They receive normal
reinforcements beginning with the second turn of the war.
\aparag The reinforcements of the Revolt of Bethlén are based only on the
provinces in \payshongrie that he controls or that are in \REVOLT . If there
are none, or if \leaderBethlen is not in play (dead or wounded), it receives
no reinforcements.

\phmil
\aparag \leaderTilly may lead any stack of \paysBaviere or its allies.
\aparag \leaderMansfeld may lead any stack of \paysPalatinat or its allies. It
can retreat with 1 \LD (only) in any neutral Protestant or mixed \HRE country
and remain there (after a battle or a retreat before battle).

\aparag[Destruction of Bohemia]\label{pIV:BR:Destruction}
If \villePrague is captured, \paysBoheme is destroyed at the end of the
current round. All its provinces are now owned by \HAB.  Its military forces
are disbanded and its provinces not yet military controlled by \HAB are
considered controlled by rebels (use Control markers of \paysBoheme); they
surrender as soon as an \ARMY \faceplus besieges them, or by regular siege
with smaller forces.

\aparag[Bethlén]
\bparag The forces of \leaderBethlen are always in restricted supply in the
national provinces of \paysHongrie (provinces with Hungarian shield). They use
the \REVOLT in \paysHongrie and the cities they control as regular supply
sources.
\bparag A force lead by \leaderBethlen can withdraw in \paysTransylvanie or
national provinces of \payshongrie owned by \TUR (by retreat or movement). If
he retreats there, he must stay there until the end of turn but may go out on
any following turn.
\bparag If \leaderBethlen and/or its forces are in \paysTransylvanie
or national provinces of \payshongrie owned by \TUR, \TUR may make a foreign
intervention against both the revolt of \leaderBethlen and \HAB. Declaring the
intervention cost 1\STAB to \TUR.
% \bparag As long as \leaderBethlen and/or its forces are in \paysTransylvanie
% or national provinces of \payshongrie owned by \TUR, \TUR has a free \CB
% against \leaderBethlen. If \TUR uses it, the forces of \leaderBethlen are
% also at war against \TUR and \TUR may enter, besiege and attack any force in
% national provinces of \payshongrie.

\phpaix
\aparag Before any peace is made, roll a test for the possible breakout of a
Religious war, according to \ref{pIV:TYW}.  A \bonus{-2} is applied to this
roll.
\aparag Add a Revolt \facemoins in a national province of \payshongrie if
\leaderBethlen is therein.
\aparag[Survival of \paysBoheme]\label{pIV:BR:Survival} If no such war occurs
and \villeVienne is controlled by the enemies of \HAB, the war end as a
victory of \paysBoheme.  The minor country is fully recreated ; \HAB has a
mandatory peace for 3 turns with \paysBoheme. \HAB loses {\bf 1} \STAB and 30
\PV ; The controller of \paysBoheme gains 30 \PV. \HAB gains the permanent
right to make the complete conquest of \paysBoheme.  \paysBoheme and
\paysPalatinat are placed in \AM of their controlling \MAJ.
\aparag \HAB and \paysBoheme stop war only when \paysBoheme is destroyed or if
\villeVienne is occupied by enemies.  Other countries use normal peace rules
(but are allied to \HAB and \paysBoheme and subjects to Separate Peace
modifiers if any).
\aparag[Victory conditions if the war becomes the TYW]
\bparag if the \nameref{pIV:TYW:Peace Prague} favours the \alliance and they
control \villeVienne, apply \ref{pIV:BR:Survival} if \paysBoheme still exists.
\bparag If the \nameref{pIV:TYW:Peace Prague} favours the \ligue, \paysBoheme
is destroyed as in \ref{pIV:BR:Destruction}.
\bparag Else, \paysBoheme remains at war after the \nameref{pIV:TYW:Peace
  Prague} and will survive the \nameref{pIV:TYW:Peace Westphalie} if not
destroyed before that during the war.

\tour{Turn 2 and following}

\phdipl
\aparag If this event does not evolve in \xnameref{pIV:TYW} (because there has
already been one, or an Appeasement of the religious fight was obtained), the
controller of \paysBoheme and \paysPalatinat may make a full intervention in
the war.



\event{pIV:Augsburg Revocation}{IV-1 (2)}{Revocation of the Truce of
  Augsburg}{1}{PBNew}

\history{Alternative history}

\condition{Check the conditions in the given order until one is found true.}
\aparag If events \xref{pIV:Bohemian Revolt}, \xref{pIV:Unity HRE} or
\xref{pIV:TYW} are happening now, do not mark off and re-roll.
\aparag If there is a \GE, apply the \xnameref{pIV:Augsburg:GE vs Northern
  Alliance}.
\aparag If \eventref{pIV:TYW} has not yet happened, apply the
\xnameref{pIV:Augsburg:HRE vs Augsburg}.
\aparag Else, apply \xnameref{pIV:Augsburg:Troubles HRE}.


\subevent[pIV:Augsburg:GE vs Northern Alliance]{Revolt of a Northern Alliance}

\phevnt
\aparag A Northern Alliance of countries of the \HRE is created.  The
countries \paysOldenburg, \paysHanovre, \paysHesse, \paysHanse, and \paysBerg
breaks free from the \GE and are allied.
\aparag \GE and \hab declare war to all those countries (and are controlled by
\SPA). The \GE is in Civil War (see \ref{chDiplo:Religious Civil War}).
\bparag \AUS has a free \CB against the whole Northern Alliance (to be used
immediately, or forfeited at the cost of {\bf 2} \STAB).
\aparag The Northern Alliance is controlled by the first Protestant \MAJ in
the list that accepts the alliance: \HOL, \ENG, \SUE, \FRA, \POL.  It has a
\CB to enter the war. If it declines, the next country in the list has the
opportunity to do the same (and so on).  The \MAJ gains all the \MIN of the
Alliance in \EG.  If nobody enters the war, \SUE controls the Northern
Alliance (but is not involved and does not gain diplomatic control).
\aparag If \HOL controls the Northern Alliance, it gains the advantages of
\ref{pIV:TYW:Northern HRE Alliance}, as long as the Alliance exists.

\phdipl
\aparag \SPA can make a full intervention as an ally of \HAB.
\aparag If they have not declined control of the Alliance, \FRA and \ENG (if
they are not Counter-Reformation) and \SUE can make a limited intervention in
the war alongside the Northern alliance.

\phpaix
\aparag If no \MAJ entered the war to control the Northern Alliance, it is
dealt with as one country for the peace in this war (except attempts of
Separate Peace), with a malus of \bonus{-4} to make peace.
\aparag A peace of level 3 or higher against the \MAJ in control (or the
Northern Alliance itself if there is none) would dissolve the Alliance in
addition to the peace.
\aparag If the war ends and the Alliance is not dissolved:
\bparag The \MIN are now normal independent countries that are no more part of
the \GE.
\bparag If the \MAJ was \HOL, it gains the benefits of \ref{pIV:TYW:Northern
  HRE Alliance}. Otherwise, the Alliance is dissolved for game purpose.
\aparag Remember that, according to \ref{pIV:TYW:German Empire}, a peace of
level 3 or higher against the \MAJHAB may dissolve the \GE.  Conversely, any
Unconditional Peace against a country once part of the \GE forces is back in
the \GE.


\subevent[pIV:Augsburg:HRE vs Augsburg]{War of Revocation of the Truce of
  Augsburg}

\phevnt
\aparag The Emperor of the HRE has the possibility to revoke this Truce (even
if it was not given in game terms). If he declines to do so, his country loses
{\bf 2 \STAB} and the event terminates.  If the Truce of Augsburg is revoked,
alliances are created in the \HRE and the \HRE is in Civil War.
\aparag[Northern Alliance] If a Northern Alliance already exists, skip this
paragraph.
\bparag A Northern Alliance of countries of the \HRE is created.  The
countries are \paysOldenburg, \paysHanovre, \paysHesse, \paysHanse, and
\paysBerg (if they exist). If there was no \terme{Truce of Augsburg} at the
beginning of the event, \pays{Hesse} and \pays{Berg} are not in the Alliance.
\bparag The Northern Alliance is controlled by the first Protestant \MAJ in
the list that accepts the alliance: \HOL, \ENG, \SUE, \FRA, \POL.  It has a
\CB to enter the war. If it declines, the next country in the list has the
opportunity to do the same (and so on).  The \MAJ gains all the \MIN of the
Alliance at \EW.  If nobody enters the war, \SUE controls the Northern
Alliance (but is not involved and does not gain diplomatic control).
\bparag If \HOL controls the Northern Alliance, it gains the advantages of
\ref{pIV:TYW:Northern HRE Alliance}, as long as the Alliance exists.
\aparag[Southern Alliance] If a Southern alliance already exists, skip this
paragraph.
\bparag A Southern HRE Alliance is created by association of \paysBaviere,
\paysMayence, \paysAlsace, \paysBade and \paysWurtemberg (if they exist).
\bparag The Southern Alliance is controlled by the first Catholic \MAJ in the
list that accepts the alliance: \AUS, \SPA, \POL, \FRA.  It has a \CB to enter
the war. If it declines, the next country in the list has the opportunity to
do the same (and so on).  The \MAJ gains all the \MIN of the Alliance at \EW.
If nobody enters the war, \MAJHAB controls the Southern Alliance (but is not
involved and does not gain diplomatic control).
\bparag \AUSMin gains the \MIN on its track, not \SPA, if \SPA accepts the
alliance.
\bparag If \MAJHAB controls the Southern Alliance, it gains the advantages of
\ref{pIV:TYW:Southern HRE Alliance}, as long as the Alliance exists.
\aparag Both Alliances are at war against each other.  The \HRE is in Civil
War.

\phdipl
\aparag \SPA can make a limited or full intervention alongside the Southern
Alliance (excepted if it declined the control and involvement).
\aparag \SUE can make a limited intervention alongside the Northern Alliance
(excepted if it declined the control and involvement).

\phpaix
\aparag A test to begin a Religious War in HRE is made at the end of the first
turn of this war, with a \bonus{-4} modifier.  See \ref{pIV:TYW} for the
result of the test and the possibility of this Religious War, and the renewal
or not of the test on following turns.  If no such war occurs, peace can be
made as usual.
\aparag{The alliances after the war}
\bparag If \HOL was controlling the Northern Alliance, the Alliance may last
after the war at the condition of \ref{pIV:TYW:Northern HRE Alliance}.
\bparag If \MAJHAB was controlling the Southern Alliance, the Alliance may
last after the war at the condition of \ref{pIV:TYW:Southern HRE Alliance}.
\bparag In other cases, the Alliances would not last after the end of the war.


\subevent[pIV:Augsburg:Troubles HRE]{Troubles in the Holy Roman Empire}

\condition{This event may happen twice, once because of \xnameref{pIV:Augsburg
    Revocation} and another time because of \xnameref{pIV:Unity HRE}}

\phevnt
\aparag \AUS, \HOL and \HIS rolls for one \REVOLT .
\aparag The effect of a diplomatic event is made on every minor country that
is part of the \HRE (fidelity/religion):
\FEforeachlist{listofminorsHRE}{\pays{\loopitem}
  (\theminorfid{\loopitem}/\theminorreligionshort{\loopitem})}{, }.



\event{pIV:Olivares}{IV-2 (1)}{Olivares}{1}{Risto}

\history{1621-1643}
\dure{as long as \strongministre{Olivares} remains the excellent minister}

\condition{}
\aparag \SPA can refuse this event if it so wishes. In that case mark-off as
played.
\aparag \SPA can freely remove \ministre{Olivares} from office at the end of
any following monarch survival phase and the event terminates.

\phevnt
\aparag \SPA receives the excellent minister \ministre{Olivares}, with values
8/9/7.  These minister values supersede the current values of the Monarch (if
they are inferior). This Minister will last for a random length of Excellent
Minister, see \ref{eco:Excellent Minister}.
\aparag If \SPA monarch dies while the this event is still in effect, use the
minister values as a basis for rolling for the values of the new
monarch. Otherwise the monarch returns with its original values when the
minister dies and play continues normally.
\aparag \SPA may receive an additional Art manufacture of level {\bf 2} (if
available, and if \SPA wants so) placed according to normal rules, and also 2
additional \TradeFLEET levels placed in any eligible trade zone (even if it
had no \SPA commercial fleet counter before, and may be in different zones).
\bparag \SPA may now move the \RES{Cloth} \MNU without any drawback
(see~\ref{chSpecific:Spain:Cloth}).
\aparag The \ctz{Espagne} can no more be attacked by Pirates and Privateers in
the Mediterranean Sea. Attacks are to be made from the Atlantic.
\aparag The malus for foreign occupation for Stability improvement is changed
from \bonus{-3} to \bonus{-5} in national provinces only, and none for other
provinces (normal rule).

\aparag The reference level of \paysGenes in \CTZ \SPA is reduced to 0 if the
Spanish player chooses so.
\aparag From now on, \SPA can raise a second privateer that can go in any \STZ
of the \CCs{Atlantic} (in Europe or in the \ROTW).



\event{pIV:Unity HRE}{IV-2 (2)}{War for the Unity of the HRE}{1}{PBNew}

\history{alternative history}

\condition{Check the conditions in the given order until one is found true.}
\aparag If events \xref{pIV:Bohemian Revolt}, \xref{pIV:Augsburg Revocation}
or \xref{pIV:TYW} are happening now, do not mark off and re-roll.
\aparag If \ref{pIV:TYW} finished during the current period, mark off and roll
for one \REVOLT in each of the following countries: \AUT, and \FRA.
\aparag If there is a \GE, apply the \xnameref{pIV:HRE:GE vs Prussian
  Alliance}.
\aparag If \ref{pIV:TYW} never happened, apply the
\xnameref{pIV:HRE:Brandenburg vs Bavaria}.
\aparag Else, use \xnameref{pIV:Augsburg:Troubles HRE}.


\subevent[pIV:HRE:GE vs Prussian Alliance]{Revolt of Brandenburg and allies}

\phevnt
\aparag A Northern Alliance of countries of the \HRE is created. \PRUMin,
\paysSaxe and \payspalatinat are created anew and break free from the
\GE. They are allied.
\aparag \GE and \hab declare war to all those countries (and are controlled by
\SPA). The \GE is in Civil War.
\bparag \AUS has a free \CB against the whole Northern Alliance (to be used
immediately, or forfeited at the cost of {\bf 2} \STAB).
\aparag This countries are controlled by the first Protestant \MAJ in the list
that accepts the alliance: \SUE, \ENG, \HOL, \FRA, \POL.  It has a \CB to
enter the war. If it declines, the next country in the list has the
opportunity to do the same (and so on).  The \MAJ gains all the \MIN of the
Alliance in \EG.  If nobody enters the war, \SUE controls the Northern
Alliance (but is not involved).

\phdipl
\aparag \SPA can make a full intervention as an ally of \HAB.
% (JCD) What about HOL?
\aparag If they have not declined control of the Alliance, \FRA and \ENG (if
they are not Counter-Reformation) and \SUE can make a limited intervention in
the war alongside the Northern alliance.

\phpaix
\aparag If no \MAJ entered the war to control of the \MIN involved, they are
dealt with as one country for the peace in this war (except attempts of
Separate Peace), with a malus of \bonus{-4} to make peace.
\aparag A peace of level 3 or higher against the \MAJ in control (or the
Alliance itself if there is none) would dissolve the Alliance.
\aparag If the war ends and the Alliance is not dissolved, the \MIN are now
normal separate countries that are no more part of the \GE. The Alliance is
then dissolved.
\aparag Remember that, according to \ref{pIV:TYW:German Empire}, a peace of
level 3 or higher against the \MAJHAB may dissolve the \GE.  Conversely, any
Unconditional Peace against a country once part of the \GE forces is back in
the \GE.


\subevent[pIV:HRE:Brandenburg vs Bavaria]{War between \paysBrandebourg and
  \paysBaviere}

\phevnt
\aparag \paysBrandebourg declares a war to \paysBaviere.  \paysSaxe and
\payspalatinat are allied to \paysBrandebourg and declares also a war to
\paysBaviere.
\aparag \AUSMin declares a full war against the enemies of \paysBaviere.
\bparag \AUS has instead a free \CB to enter war as an ally of \paysBaviere,
and will lose {\bf 2} \STAB if it does not use it.

\phdipl
\aparag Each \MAJ that controls one of the involved countries may react as per
the usual rules to enter in limited intervention (only).
\aparag \SPA may make a limited intervention as an ally of the side of
\paysBaviere.

\phpaix
\aparag A test to begin a Religious War in HRE is made at the end of the first
turn of this war, with a \bonus{-2} modifier.  See \ref{pIV:TYW} for the
result of the test and the possibility of this Religious War, and the renewal
or not of the test on following turns.  If no such war occurs, peace can be
made as usual.



\event{pIV:War Persia Turkey}{IV-3 (1)}{War between Turkey and
  Persia}{1}{Risto}

\history{1606-1639}

\condition{This event has the same conditions and effects as \ref{pIII:War
    Persia Turkey}. It is nonetheless a different event (thus both can happen
  separately).}



\event{pIV:Persian Safavids}{IV-3 (2)}{Persian Safavids}{1}{PB}

\history{1590-1722 -- The high tide of Shah Abbas.}

\condition{}
\aparag This event is the same as \ref{pIII:Persian Safavids}.  If it did not
happen, apply immediately its effects. Apply additionally \ref{pIV:Persian
  Safavids:Fall Ormus}.
\aparag If it happened and main provinces of \paysperse are conquered,
activate a Persian Uprising as per the rules.
\aparag Else, apply the following events.

\phevnt
\aparag[Fall of Ormus]\label{pIV:Persian Safavids:Fall Ormus} \dipAT and
\dipFR status with \paysOrmus are immediately broken to \dipNR. This might
cause an Activation of \paysperse against a \TP in \paysOrmus.
\aparag[Conquest of Oman] \paysperse attacks \paysOman and that results in
breaking all diplomatic status of \paysOman. This applies also to military
\dipAT imposed by \PORmin (troops are redeployed).
\aparag[Submission of \granderegion{Afghanistan}]
\bparag \granderegion{Afghanistan} is no more part of the \paysMogol or
\paysAfghans (which is destroyed at this point), but submitted to
\paysperse. The Natives are used by \paysperse in this region.
\bparag As long as \paysPerse masters \granderegion{Afghanistan}, the Silk
resources of this region may be exploited by \provinceOrmus (usual concurrence
with \TP or \COL in \granderegion{Afghanistan}).
\bparag Persian units can go in \granderegion{Afghanistan} and have supply in
every provinces. But only \villeHerat and the European provinces of \paysperse
are supply sources.
\bparag \RUS and \TUR have \OCB against \paysperse as long as it owns
\granderegion{Afghanistan}.
\bparag \granderegion{Afghanistan} can be conquered later by \paysMogol, or
can become \paysAfghans again by \ref{pVI:India:Afghan Empire}.
\bparag \paysperse also loses the area in a losing Peace of level 2 or higher
(in regular or Overseas war) that has no other condition of peace. In Overseas
Wars, the occupation of a province without city counts as a province
occupied. In every war, the control of \villeHerat and its province counts as
for a Persian province.
\bparag When \paysperse loses the area, all the effects described here are
nullified.



\event{pIV:Portuguese Revolt}{IV-4 (1)}{National Revolt of the
  Portugal}{1}{Risto}

\begin{todo}
  Province Tanger should go to Morocco. Helper of POR should gain a predisio
  on Tanger + a \TP of POR in case of victory (no reannexion). HIS should not
  be able to attack if at war otherwise. Helper should be first in order at
  Methuen. Helper should be Catholic?

  Maybe swap Portuguese revolt with Alaouite dynasty and re-add Portuguese
  revolt as secondary event in pV (typically of WoSS which is four times in
  the table). The real war only started in 1660, the turning point between pIV
  and pV.
\end{todo}

\history{1640-1668}

\condition{Occurs only if \xnameref{pIII:POR Ann.:Portugal Annexed} is
  currently in effect.}
\aparag Else, if \ref{pIII:Portuguese Annexation} was never rolled for, do not
mark off and re-roll.
\aparag Otherwise treat as a \RD instead, with a \REVOLT in \SPA.

\phevnt
\aparag All effects of the \xnameref{pIII:POR Ann.:Portugal Annexed} are
cancelled and \paysPortugal returns to play as a minor country. \paysPortugal
receives the same provinces it had at the time of its annexation to \SPA
notwithstanding who currently owns such provinces. It also receives all
Portuguese \COL/\TP, missions, forts/fortresses, commercial fleets etc. that
are currently in Spanish hands.
% (Jym) Included \provinceTanger and other provinces given to SPA during
% annexation (JCD) Yes, why all of the provinces?
\aparag All non-Portuguese \COL in \continent{Brazil} receive a \REVOLT
\faceplus controlled by \paysPortugal. They can't extend outside the regions
of \continent{Brazil}.
\aparag All non-Portuguese troops inside its territories are removed as per
normal peace phase.
\aparag All Portuguese troops are removed from the map, as \paysPortugal is
initially at peace (keep the basic forces in the \ROTW where needed).
\aparag \ENG may accept \paysportugal in \EG; if \ENG declines, same to \FRA,
then to \SUE; if no country accepts, it remains neutral.

\phdipl
\aparag All players who are forced to cede provinces to Portugal by this event
receive a temporary free \CB to be used this turn.
\aparag Players who want to fight against Portuguese \REVOLT in their own \COL
have to declare an Overseas war against \paysPortugal and have a free \OCB to
do so.  Else, their \COL is freely given to \paysportugal (no loss of \STAB or
\VPs).
\aparag \SPA receives a free \CB that lasts until the end of the next period
and can be used multiple times.

\phpaix
\aparag The Portuguese Revolt in a \COL causes the loss of at most {\bf 1}
\STAB to each \MAJ.
\aparag Any \COL having 2 Portuguese \REVOLT \faceplus in it is immediately
annexed by \paysPortugal.
\aparag If \SPA uses its free \CB against \paysPortugal and wins an enforced
unconditional surrender over Portugal, it can reapply \xnameref{pIII:POR
  Ann.:Portugal Annexed}. All Portuguese possessions as they are now are
reannexed to \SPA as described there. Reannexation of Portugal as by
\numberref{pIII:POR Ann.:Portugal Annexed} is only possible in wars \SPA
started using its free \CB. In addition, \SPA gains a \Presidio in
\provinceTanger if the province was Portuguese.
\aparag If \SPA uses its free \CB against \paysPortugal, but does not win an
enforced unconditional surrender over it, the controller of \paysPortugal
receives 30 \VP when peace is made. This can occur several times.
refshort{pIII:POR Ann.:Portugal Annexed} is only possible in wars \SPA started
using its free \CB.  + \fphase In addition, \SPA gains a \Presidio in
\province{Tanger} if + the province was Portuguese.  \ephase If \SPA does use
its free \CB against \pays{Portugal}, but does not win an enforced
unconditional surrender over it, the controller of \pays{Portugal} receives 30
\VP when peace is made. This can occur several times.
\aparag Whatever the result of the war, if \ENG was intervening in the war, it
gains \provinceTanger if the province was Portuguese.



\event{pIV:Morocco}{IV-4 (2)}{Alaouite dynasty in
  \paysmaroc}{1}{PB+JymNew}

\begin{todo}
  Maybe here for giving back Tangier (except presidio?) to Morocco?
\end{todo}

\history{1631}
\dure{Until the end of the game}

\effetlong
\aparag \TUR has a malus of \bonus{-3} to diplomacy with \paysmaroc.
\aparag \paysmaroc loses its \corsaire counter.
\aparag Fidelity of \paysmaroc is now 10.



\event{pIV:Act Navigation}{IV-5}{Act of Navigation}{1}{RistoMod}

\history{1651}
\dure{until English defeat in a war caused by this event, or by event
  \ref{pV:Glorious Revolution}}

\condition{}
\aparag Can occur only if \ENG is currently \PROTANG. Otherwise re-roll.
\aparag Can occur only if \ref{pIV:English Civil War} has already occurred
(not necessarily ended). Otherwise re-roll.
\aparag \ENG can refuse the event, in which case it is marked off and \RD is
applied instead.

\phevnt
\aparag All non-English commercial fleet counters in \ctz{Angleterre} are
eliminated and \ENG receives 2 \TradeFLEET levels in \ctz{Angleterre} (up to 6
levels). All powers that lose their counters as a result of this, receive a
\CB against England until the end of current period.
\aparag From now on, only \ENG can place \TradeFLEET levels in
\ctz{Angleterre}.
\aparag From now on, all \MAJ have an \OCB against \ENG, usable once each
period.

\phadm
\aparag \ENG may ignore restriction of~\ref{chAdministration:Pioneering} for
this turn.

\phpaix
\aparag If a \CB against \ENG received through this event has been used by a
player, and if such a player wins at least a level 4 victory against \ENG, he
may renounce the effects of this event instead of any other peace conditions
(all the allies must agree with this as per normal peace procedure).
\aparag In such a case all non-English \TradeFLEET levels in \ctz{Angleterre}
lost due to this event are returned and \ENG \TradeFLEET in \ctz{Angleterre}
is reduced to 1 whatever its current level.



\event{pIV:Union Scotland}{IV-6}{Personal Union between England and
  Scotland}{1}{Risto}

\history{1603}
\dure{until \ref{pVI:Act Union} occurs or the union is dismantled by
  \ref{pIV:English Civil War}}

\condition{Can occur only if \paysecosse is at peace with \ENG.  Otherwise
  re-roll and do not mark off as played.}

\phevnt
\aparag \paysecosse becomes a permanent \VASSAL of \ENG whatever its current
situation.
\aparag If \paysecosse is currently at war, its opponent must immediately
either accept a white peace with it, or declare war to \ENG with a free \CB.
Normal call for allies can be made for such a war at this point.



\event{pIV:English Civil War}{IV-7 (1)}{English Civil War}{1}{RistoMod}

\history{1642-1648}

\condition{}
\aparag If \monarque{Elisabeth I} rules in \ENG, do not mark off and reroll.
\aparag If \ENG is \CATHCR or \CATHCO and period is III, do not mark off and
reroll.
\aparag If \ENG is currently at war, it offers an immediate white Peace or
Armistice to all its enemies, and will renew the offer at the end of each
turn.
\bparag This event is activated as soon as \ENG is at peace or in armistice
with every other \MAJ (\MIN automatically accept an Armistice).

\phevnt
\aparag A Religious Civil War (\ref{chDiplo:Religious Civil War}) erupts in
\ENG between the \parl and the \paysRoyalists.

\aparag[Which side is played ?]
\bparag If \ENG is \CATHCR by choice in \ref{pI:Reformation2}, the player
keeps playing \royal.
\bparag If \ENG is \CATHCO, or \CATHCR by forced conversion, the player
chooses the side he will play.
\bparag If \ENG is Anglican or Protestant, the player plays \parl.
\bparag The other side will be the Rebels; the \MAJ controlling the Rebels
will be called \REB.
\bparag\label{pIV:ECW:Monarchs} The \royal are governed by the English Monarch
before the event (and he can be used as a general). The \parl are ruled by a
Monarch \monarque{Parliament} of values 5/8/8 that makes no test of Survival.
It gives a bonus of \bonus{+2} to the rolls for all administrative actions
(except exceptional taxes). It may not be used as a general.

\aparag If not played by \ENG, \parl is played by the first Protestant country
in the list: \HOL, \FRA, \SUE or else by \POL.

\aparag Three \REVOLT are rolled for in \ENG. These \REVOLT are hostile to
both sides and controlled by \TUR.

\aparag If not played by \ENG, \royal are played by the first \CATHCR \MAJ in
the list: \SPA, \FRA, \HOL, \VEN else by the first Catholic \MAJ in the list:
\SPA, \FRA, \VEN, \SUE, \POR, \POL.  Failing that it is played by \RUS.

\aparag[Initial position]
\bparag \royal control \provinceMidlands, \provinceCornwall, \provinceDurham
and 1d10/3 (round down) provinces adjacent to \provinceMidlands (to be chosen
by their controller). Add \bonus{+2} to the roll if \ENG was
Counter-Reformation or Protestant. \royal controls all (non-revolted)
provinces in \regionIrlande.
\bparag The \parl control all other (non-revolted) provinces in \ENG.
\bparag \royal and the \parl receive up to the equivalent of basic land forces
of \ENG; the Rebels take the forces before (so they can take everything is
there is not enough). Additional forces are removed.
\bparag The Rebels add 1\LD (Veteran) in any controlled province, and 1\LD
(Conscript) in \provinceDurham (if \royal) or \provinceWessex (if \parl).
\bparag \ENG loses 1 \ND, and the rest is controlled by the \parl.
\bparag All named leaders in play are controlled by the \parl.

\aparag[Economic consequences]
\bparag \ENG loses one-third of its treasury, and at least 50\ducats (this
might cause a Bankruptcy).
\bparag Two \PIRATE\faceplus are placed in \CTZ England.
\bparag All \TP, \COL, \TradeFLEET, etc., remain under the control of \ENG.

\aparag If \ref{pIV:Union Scotland} is in effect, apply \xnameref{pIV:ECW:War
  Scotland} in addition.

\phdipl
\aparag If \ENG was \CATHCR, \SPA if also \CATHCR may make a full intervention
on the side of the \royal.
% (JCD) I added /Anglican there.
\aparag If \ENG was \PROTANG, \SPA if \CATHCR may make a limited intervention
on the side of the \royal.
\aparag If \ENG was not \CATHCR, \HOL if Protestant may make a limited
intervention on the side of the \parl.

\phadm
\aparag[Reinforcements]
\bparag The Rebels roll for reinforcement with offensive status, or naval
status at \bonus{-3}, during all the war. On the first turn, they roll for
offensive with a modifier of \bonus{+4} if \ENG was Protestant, \bonus{+2} if
it was \CATHCR or \PROTANG, of \bonus{0} otherwise (\CATHCO).
\bparag On following turns, they receive a modifier of \bonus{+1} for every 2
provinces they control, with a maximum of \bonus{+4}.
\bparag If the Rebels are the \parl, they can take up to 2 \LD as \ND instead.
\bparag The Rebels have as many counters as \ENG available.
\aparag \ENG uses normal purchase rules, except that its purchase limits are
doubled during the Civil War.
\aparag The \royal receive the general \leaderwithdata{Rupertroy} on the first
turn of the war; he will last 7 turns.% He has a double usage as an admiral.
\aparag The \parl receive the general \leaderwithdata{Cromwell} at the end of
the first turn of the war (before the Peace Segment).  He will last for the 5
following turns. The \parl benefits from a Military Revolution at that point
(\textit{The New Model Army}, see rules \ruleref{chAdministration:Military
  Revolutions}, that is to take immediately any Land Technology available at
most in 4 turns, and in the mean time, is raised to the highest Technology
available at that time).
\aparag The \royal have the Land Technology of \ENG at the beginning of the
event. If played by \ENG, they may raise their technology as per usual rules;
else their Land technology is raised by {\bf 1} each turn of the war beginning
with the second.

\phpaix
\aparag The Civil War ends only if either party achieves both following
conditions:
\bparag Military control of \province{East Anglia} and five other English
National provinces with at least 3 ports.
\bparag Elimination of all enemy army counters, or at least two major
victories against them.
\aparag If the \royal win, \ENG is ruled by its previous Monarch and becomes
\CATHCR (exception: if \ENG was \CATHCO, it remains so). \leaderRupertroy is
kept as a general; land technology of \ENG is at the level reached by the
\royal.
% (JCD): I added (or remains)
\aparag If the \parl wins and \ENG was Catholic or \PROTANG, \ENG becomes (or
remains) \PROTANG.  It is ruled by the \monarque{Parliament}
(see~\ref{pIV:ECW:Monarchs}).
\bparag If \leaderCromwell is in play at the end of the war, it becomes Lord
Protector of the Kingdom, and is an English Monarch that raises the values of
the \monarque{Parliament} to 8/8/9. His Reign is to last the number of turns
remaining for the general (of the initial 5 turns).  A test of survival has to
be done for him. As long as his reign continues, \ENG gains a free maintenance
of one \ARMY\faceplus.
\bparag When \leaderCromwell dies, or at the beginning of the sixth turn after
the end of the Civil War, apply \ref{pIV:English Restoration} as one of the
event of the turn.
\bparag \leaderRupertang becomes an admiral only, kept by \ENG as one of its
own.

\aparag If \ENG was Protestant (not \PROTANG) and the \parl wins, \ENG remains
so. It is rules by the \monarque{Parliament} (see~\ref{pIV:ECW:Monarchs}).
\bparag If \leaderCromwell is in play at the end of the war, he stays as a
general only. \leaderRupertang is not used by \ENG.
\bparag At the beginning of the sixth turn after the end of the Civil War,
roll for a new Monarch on columns 9 for the three values. \ENG is ruled by a
Protestant Republic lead by some strong Lord Protector of the Commonwealth
(represented by the Monarch).
\aparag Regardless of the winner, \leaderMonck and \leaderBlake are admirals
from now on.


\subevent[pIV:ECW:War Scotland]{War with Scotland}

\phevnt
\aparag \paysecosse declares war against the \royal and becomes neutral.
\paysecosse is controlled by \FRA, but no allies can ever take part in this
war. This declaration of war does not trigger a truce in the civil war as per
normal rules.

\phadm
\aparag \paysecosse rolls for reinforcements in offensive status.  It has a
minor general added to its base forces.

\phmil
\aparag Scottish units may not enter England during the first 2 rounds of
their war.
% (Jym): What? (Pierre) I do insist
\aparag On the turn following their entrance in England, \royal gain as added
reinforcements \leaderwithdata{Montrose}, 2 \LD and control of one mountainous
province in \paysecosse of their choice.

\phpaix
\aparag When the Civil War ends, \ENG may decide to continue an on-going war
against \paysecosse (it will be counted as the second turn of the war).
\aparag If \ENG (\royal or, after the end of the Civil War, the \parl) scores
an enforced unconditional victory over \paysecosse during this war, Scotland
is restored to permanent \VASSAL of \ENG as per \ref{pIV:Union Scotland}.  In
all other cases, it reverts to a normal minor.



\event{pIV:English Restoration}{IV-7 (2)}{The Parliament and the English
  Kings}{1}{PBNew}

\history{1660}

\condition{May not happen if the \xnameref{pIV:English Civil War} is not
  finished yet. Re-roll and do not mark off.}

\phevnt
\aparag If \ENG is \PROTANG or \CATHCO, apply \xnameref{pIV:ENG
  Restoration:Restoration}.
\aparag If \ENG is \CATHCR, apply \xnameref{pIV:ENG Restoration:Asking
  Reforms}.
\aparag If \ENG is Protestant, apply \xnameref{pIV:ENG Restoration:War Against
  Puritanism}.


\subevent[pIV:ENG Restoration:Restoration]{The Restoration of the English
  Kings}

\phevnt
\aparag \ENG has the choice of crowning now the Pretender (an Heir of the
Monarch overthrown by \ref{pIV:English Civil War}); if not, \ref{pV:Glorious
  Revolution} is applied now (with worsened consequences).
\aparag If the Pretender is crowned, roll for his values using those of the
Monarch overthrown by the \nameref{pIV:English Civil War}. The effects of
\monarque{Cromwell} or the \monarque{Parliament} are ended (and
\leaderCromwell is put out of play).
\aparag \ENG receives the general (also usable as admiral) \leader{Duke of
  York} that will stay for 5 turns (note: he actually became king in 1685 but
we choose to ignore this and separate the general from the king).


\subevent[pIV:ENG Restoration:Asking Reforms]{The Parliament asks for more
  reforms}

\phevnt
\aparag \ENG has to choose one of the 2 following attitude.
\aparag[Reforms granted]
\bparag \ENG becomes Anglican. It loses {\bf 2} \STAB and rolls for 2 \REVOLT
.
\bparag \SPA, if \CATHCR, has a free \CB against \ENG.
\aparag[Refusal]
\begin{todo}
  ``CHANGE'' (Pierre's notes).
\end{todo}
\bparag Apply \ref{pV:Glorious Revolution} now (with worsened consequences).


\subevent[pIV:ENG Restoration:War Against Puritanism]{Civil War between
  Protestants and Puritans}

\phevnt
\aparag \ENG is now in Civil war (\ref{chDiplo:Religious Civil War}) between
two sides: the (Puritans and Calvinist) Rebels (possibly with Orange
Partisans) and the (Protestant) Royalists. Catholics rebel against both sides.
\bparag The Rebels are controlled by a Protestant \FRA, or \HOL (or \SUE if
there is no \HOL). They use the \paysroyalistes counters.
\bparag The (Protestant) Royalists are controlled by \ENG and use \ENG
counters; all \ENG leaders are with them.
\aparag Four Rebel \REVOLT are rolled for in England (re-roll until in English
owned territory). They control all the fortresses.
\bparag A Rebel \ARMY \facemoins and a \LeaderG are placed in one of these
provinces.
\aparag Catholic \REVOLT \faceplus are placed both in \provinceConnacht and
\provinceMumhan and the \REVOLT control both fortresses.

\phdipl
\aparag The controller of the Rebels have a \CB against \ENG to make a limited
intervention against \ENG this turn, that can become a full intervention on
the second turn.

\phadm
\aparag The Rebels roll for reinforcements in offensive or naval status (but
with \bonus{-2} for naval).
\aparag All reinforcements must be placed in a province with existing rebel
units, allied units, or controlled cities (\REVOLT are not enough). If none,
no reinforcements are received.

\phpaix
\aparag Peace is determined with usual rules except that:
\bparag The Rebels surrender unconditionally if they have no forces nor
\REVOLT left (fortresses do not count).
\bparag If the English King is overthrowned by \REVOLT, it also surrenders
unconditionally to the Rebels and their controller.
\aparag If the Rebels win, \ENG will have a \terme{Dynastic Crisis} next turn,
and loses 50 \PV.
\bparag \eventref{pVI:Act Union} is broken. If it did not happen yet, may
occur later.
\aparag If the Rebels win unconditionally and their controller was involved in
full intervention, additional consequences are:
\bparag \ENG makes a mandatory Dynastic Alliance with the controller of the
Rebels and must give a \COL or \TP as dowry.
\bparag \ENG makes a mandatory offensive alliance with the controller of the
Rebels for 2 turns. It cannot declare war against it (except with \CB from
events; then, the alliance has to be broken with the usual cost in \STAB).
\bparag \eventref{pVI:Act Union} is broken. If it did not happen yet, it may
not occur later.



\event{pIV:London Stock Exchange}{IV-8 (1)}{Creation of the London Stock
  Exchange}{1}{Risto}

\history{1554}

\condition{\ENG chooses to apply this event or \ref{pIII:East Indian
    Company}. Mark the one that is chosen.}

\phadm
\aparag \ENG may ignore restriction of~\ref{chAdministration:Pioneering} for
this turn.

\effetlong
\aparag \ENG can from now on lend 250\ducats per turn to other countries.
% in the Diplomacy phase, plus 100\ducats during the turn (instead of
% 100\ducats plus 50\ducats).

\aparag From now on, \ANG receives a bonus for its International Loan rolls
and Bankruptcy rolls.

\aparag From now on, \ANG is more resilient to exceeding limits in \MNU.

% \begin{oldcompta}
%   \aparag From now on \ENG receives a bonus equal to its \DTI to all die-rolls
%   on international loan amount and interest (not length) in the loans table
%   % (Jym) Table just said "international" and usually more up to date
%   \aparag \ENG is also more resistant to Bankrupt and more tolerant to
%   trespassing of commercial limits.
%   % (Jym) According to the rule of chapter 3, it is actually less resilient
% \end{oldcompta}



\event{pIV:Amsterdam Stock Exchange}{IV-8 (2)}{Creation of the Amsterdam Stock
  Exchange}{1}{Risto}

\history{1608}

\condition{This event is the same as \ref{pIII:Amsterdam Stock Exchange}.}



\event{pIV:Dutch Colonial Dynamism}{IV-9}{Dutch Colonial Dynamism}{3}{Risto}

\condition{\HOL chooses to apply this event or \ref{pIII:VOC}. Mark the one
  that is chosen.}

\phevnt
\aparag \HOL receives an additional commercial fleet level to any eligible
\STZ zone in \ROTW map (even if none existed before).

\phdipl
\aparag For this turn \HOL receives a bonus of \bonus{+2} to all diplomatic
actions made on countries from the \ROTW map.

\phadm
\aparag \HOL receives an additional and free strong \TP placement attempt.
\aparag For this turn \HOL receives a bonus of \bonus{+1} to all
administrative actions made in \ROTW map.
\aparag \HOL may ignore restriction of~\ref{chAdministration:Pioneering} for
this turn.



\event{pIV:Liberum Veto}{IV-10 (1)}{Liberum Veto}{2}{PB}

\history{1652}

\phevnt
\aparag The conditions of the \textit{Liberum Veto} (see
\ref{chSpecific:Poland:Liberum Veto}) are now enforced.

\phadm
\aparag If \POL is at peace after the diplomacy phase of this turn, it loses 2
in \STAB.

\effetlong
\aparag Each time a new dynasty is elected in \POL, it can decide to impose
Absolutism in the Republic. This decision is made at the phase of the monarch
survival (before the events) ; it causes an additional event, \ref{pIV:Polish
  Civil War}. There can be no further additional event at this turn.



\event{pIV:Great Elector}{IV-11}{The Great Elector Friedrich-Wilhelm of
  Prussia}{1}{PB}

\history{1640-1688}

\phevnt
\aparag \leader{Friedrich-Wilhelm} is now the ruler of \paysBrandebourg and a
general [A.2.3.3]. He will last 6 turns.  The basic force of this country is
now one \ARMY\faceplus, 1 \LD, 1 \fortress and 1 general. Its counter limits
are 2 \ARMY and 5 \LD.  The fidelity of the country is 9 from now on.
\aparag \paysBrandebourg claims the \region{Duche de Prusse}:
\provincePreussen, \province{Ost Pommern} and \provinceMemel.
\bparag Minor countries cede those provinces immediately to \paysBrandebourg.
\bparag Major countries have the possibility to cede them or not (and lose \PV
normally).

\phdipl
\aparag If a country declares a war against a \MAJ that owns one of those
territories, he can ask for a full intervention of \paysBrandebourg as an ally
(which is put in \EW immediately).
\aparag If \POL owns provinces of the \region{Duche de Prusse}, it can cede
all of them to \paysBrandebourg in exchange for an alliance with
\paysBrandebourg. \POL does not lose the \PV.  \paysBrandebourg signs a white
peace, is put in \EW of \POL and may be called as ally by \POL in any war it
is currently involved in.
\bparag \POL is now the first power in the list of preference of
\paysBrandebourg.

\phpaix
\aparag In any war involving \paysBrandebourg, only this country may annex a
province of the Duchy of Prussia if its alliance wins; if its alliance wins,
it asks for one province or refuses the peace (so that the other powers must
break their alliance and make a separate peace).



\event{pIV:Oxenstierna}{IV-12 (1)}{Oxenstierna}{1}{PBNew}

\condition{Same event as \ref{pIII:Oxenstierna}.}



\event{pIV:Union Poland Sweden}{IV-12 (2)}{Union between \paysmajeurPologne
  and \paysmajeurSuede}{1}{PB}

\condition{Same event as \ref{pIII:Union Poland Sweden}.}



\event{pIV:Torstensson}{IV-13 (1)}{Torstensson's War}{1}{PB}

\history{1643-1645}

\phevnt
% (JCD) TODO: adapt to DANdan
\aparag \SUE has a mandatory free \CB against \paysDanemark at this turn (even
if their are allied in another war).
\aparag If \SUE refuses the \CB, it loses 2 \STAB.



\event{pIV:Swedish Nobles Unrest}{IV-13 (2)}{Agitation of the Swedish
  Nobles}{1}{PBNew}

\history{1650's}

\phevnt
\aparag If \SUE is Catholic and \ref{pIII:Religious War Sweden} did not happen
yet, it occurs now.
\aparag If \SUE is \PROTRIG, roll for two \REVOLT in \SUE.
\aparag If \SUE is \PROTTOL and at war, rolls for one \REVOLT in \SUE, \SUE
loses 2x \STAB and its monarch changes (abdication of the Queen Kristin).
\aparag If \SUE if \PROTTOL but not at war, roll for 4 \REVOLT in \SUE (do not
place the \REVOLT if not inside \SUE, but do not reroll either) and a Revolted
\ARMY appears in one of those provinces with a general.
\aparag The resulting \REVOLT are controlled by \SPA.


\event{pIV:La Rochelle}{IV-14}{Revolt of La Rochelle}{1}{RistoMod}

\history{1626}
\dure{Until the suppression of the \REVOLT in \provincePoitou and the conquest
  of \ville{La Rochelle}.}

\condition{}
\aparag If \ref{pIII:FWR} is not finished yet, do not mark off and re-roll.
\aparag If the owner of \provincePoitou is Protestant, roll on its Revolt
table and place a \REVOLT\Faceplus if this is a Catholic province, and a
\REVOLT\Facemoins otherwise. The event is marked off and considered as played.

\phevnt
\aparag Place 2 \REVOLT \faceplus and a \LD in \provincePoitou.  Retreat all
other units from the province.
\bparag Roll for two Revolts in \FRA. Place a \REVOLT\Facemoins if the
province is Protestant (or mixed if \FRA is \CATHCR) and nothing otherwise.
% \aparag Place 2 Revolts \facemoins in other protestant provinces (randomly),
% or in Protestant or Mixed provinces if \FRA is \CATHCR.
\aparag The fortress of La Rochelle is controlled by the Rebels and upgraded
to the highest level available to the owner of the province.
\aparag Place a \PIRATE\faceplus in \CTZ of \FRA.
\aparag The Rebels/\REVOLT are controlled by \ENG, or \FRA if \ENG owns the
province. This war is a Religious Civil War (see \ref{chDiplo:Religious Civil
  War}) between Protestants and Catholics and normal Foreign interventions are
allowed.

\phadm
\aparag As long as the event lasts, the owner of \provincePoitou has a malus
of \bonus{-1} to all its administrative actions in the \ROTW.

\phmil
\aparag If a Foreign power enters a land province in the power at war that is
not \provincePoitou during its intervention, it loses 1 \STAB.
\aparag If the owner of \provincePoitou is \FRA and \ministre{Richelieu} is in
the game, consider that the port of the fortress is under blockade if a french
army besieges it.

\phpaix
\aparag If the fortress is controlled by the Rebels, it counts has a \REVOLT
\facemoins for the loss of \STAB by the owner of \provincePoitou due to
\REVOLT .
\aparag The owner of \provincePoitou may cede the province to the controller
of the \REVOLT , losing 30 \PV for doing this.
\aparag The controller of the \REVOLT earns 5 \PV at the end of each turn that
the Rebels exist (\REVOLT or fortress in \provincePoitou).



\event{pIV:Richelieu}{IV-15}{Richelieu}{1}{RistoMod}

\history{1624-1642}
\dure{as long as \strongministre{Richelieu} remains the excellent minister}

\condition{}
\aparag If \ref{pIII:FWR} is not finished yet, do not mark off and re-roll.
\aparag \FRA can refuse this event if it so wishes. In that case mark-off a
played.
\aparag \FRA can freely remove \ministre{Richelieu} from office at the end of
any following monarch survival phase and the event terminates.

\phevnt
\aparag \FRA receives automatically the excellent minister
\ministre{Richelieu}, with values 9/8/7.  These minister values supersede the
current values of the Monarch (if they are inferior). This Minister will last
for a random length of Excellent Minister, see \ref{eco:Excellent Minister}.
\aparag \FRA gains one level of \TradeFLEET in any \CTZ or \STZ of its choice.

\phadm
\aparag As long as \ministre{Richelieu} lives, \FRA has a bonus of \bonus{+2}
to any die-roll for External Administrative Actions and to improve its \FTI.
\aparag \FRA may ignore restriction of~\ref{chAdministration:Pioneering} for
this turn (only).

\phinter
\aparag When \FRA monarch dies, his successor is \monarque{Louis XIV}.



\event{pIV:Fronde}{IV-16}{The Fronde}{1}{PB}

\history{1648-1653}
\dure{3 turns or as long as \monarque{Louis XIV} is not adult (whichever is
  the latest). In any case, it ends after the turn of revolts.}

\phevnt
\aparag If \monarque{Louis XIV} has not already been king of \FRA, the current
king of \FRA dies and the new king is \monarque{Louis XIV}.
\bparag \monarque{Louis XIV} has values 7/6/9, scheduled to last 12 turns and
starts as a baby.
\bparag He'll become adult at the beginning of the fourth turn of reign, thus
ending the event.
\aparag If due to \xnameref{pIV:Richelieu}, \ministre{Richelieu} was still in
charge, then during the first two turns of reign of \monarque{Louis XIV},
\ministre{Mazarin} will be minister with values 7/8/7.
\aparag If \monarque{Louis XIV} is already king, or if his reign is already
finished, then the event lasts for 2 turns.

\phdipl
\aparag Until the end of the event, \FRA may only offer a white or losing
peace to all minors, and peace based on the peace differential to each major
countries, with a maximum level of 1 in the favour of \FRA.
\bparag At each turn, \FRA offer and cannot refuse Armistices with opponents.
\bparag Neutral minor countries always accept that peace.
% \bparag Other countries (either majors or minors) choose their attitude
% freely toward that proposal: accept, refuse or armistice.
\bparag At the third turn of the event, if \ministre{Mazarin} is minister,
then major countries cannot refuse an armistice.
\aparag \FRA may not declare war as long as the event lasts (except
\xnameref{pIV:TYW} and \xnameref{pV:WoSS}).
\aparag If, at the end of a diplomacy phase, \FRA is not at war (don't count
armistices), the Fronde happens.

\tour{Turn of the revolts}

\phdipl
\aparag Half of French troops in Europe become rebel. \FRA choose a stack of
troops staying loyal, thus taking up to half the total number of \LD (rounded
down). The rest becomes the troops of the Fronde.
\bparag If in play, \leaderConde becomes leader of the Fronde. Otherwise, a
randomly chosen general (a named one if there is one in play) becomes leader
of the Fronde.
\aparag The Fronde is controlled by a country currently at armistice with
\FRA. If none exists, the order of priority to control the Fronde is: \HIS,
\ANG, \HOL, \AUS, \POL, \RUS, \SUE, \TUR.
\aparag Naval forces, admirals, everything in the \ROTW as well as
administrative counters (\MNU, \ldots) stay in the control of \FRA.

\phadm
\aparag \FRA collects neither land nor vassals income this turn. \FRA does get
other incomes.
\aparag The Fronde rolls for reinforcements with offensive attitude and no
modifier.
\aparag No side may get reinforcements such that its total force (in Europe)
is above the basic force of \FRA for the current period.

\phmil
\aparag Countries in armistice with \FRA can enter the civil war on the side
they want.
\aparag Fleet of \FRA may stay in ports controlled by the rebels without
penalties.
\aparag Except for the capital of \FRA, fortresses in France are friendly to
both sides.
\bparag A province is controlled by one side if it has an army in the province
and there is no enemy troop besieged in the fortress.
\bparag Other provinces are considered friendly to both sides for supply or
movement cost.
\aparag The capital of \FRA is always controlled by the loyalists until the
Fronde takes the fortress.

\phpaix
\aparag The side controlling the capital of \FRA at the end of turn wins.
\bparag If the Rebels win, \monarque{Louis XIV} (and \ministre{Mazarin}) is
overthrown. During the next turn, there will be a dynastic crisis in \FRA. The
player controlling the Fronde gains 10 \VP.
\bparag In any case, both the loyalist and Fronde's units become units of \FRA
as soon as the event is finished.



\event{pIV:Times of Troubles}{IV-17 (1)}{The Times of Troubles in
  Russia}{1}{PB}

\history{1605-1613}

\condition{}
\aparag If \monarque{Ivan IV} is not dead yet, do not mark-off and re-roll.
\aparag If \RUS chose \terme{Religious Tolerance}, mark off and use \RD
instead.
\aparag If \RUS is at war, the event is pending. It will activate at the
beginning of the first turn where \RUS is at peace and a roll of 6 or higher
is obtained on 1d10.

\tour{Turn 1}

\phevnt

\aparag The Russian monarch dies and is replaced by \monarque{Boris Godunov}.
His values are 5/8/4 and he will reign 5 turns; he is a general
\leaderwithdata{Godunov}.

\aparag \RUS is now in Religious Civil War (see \ref{chDiplo:Religious Civil
  War}). Rebels are Catholic; loyalists (\RUS) are Orthodox.

\aparag Roll for 6 \REVOLT in Russia. Only provinces actually in \RUS revolt,
other rolled-for are ignored. The \REVOLT are controlled by \POL.

\aparag Rebels gain one \ARMY\faceplus in one province in \REVOLT, and the
control of the city.

\aparag Rebels own any revolted province with no Russian armies in there
(except \province{Moscou}) and provinces they control. These provinces are
their supply sources.

\aparag \RUS owns all non-revolted provinces they control. They are their
supply sources.

\aparag All other provinces are disputed. Supply of both sides may cross those
provinces if there is no enemy force within.

\phdipl

\aparag During the event, \RUS may ask for help of \SUE. The condition is the
cession of one Russian province to \SUE; if this province is revolted, it
becomes Swedish only when it is no more revolted and its is controlled by \RUS
or \SUE. During the rest of the event, this province (even Swedish) can be
entered and attacked by all belligerents.

\bparag If an intervention of \SUE is agreed upon, \SUE has to commit at least
4 \LD in Russia, following the conditions of limited intervention.  \SUE can
not withdraw any force sent in Russia.

\aparag Major countries may make \terme{Foreign Intervention} in this war.

\phadm

\aparag Rebels receive offensive reinforcements at each turn, using the
provinces they own.

\aparag Rebels have the general \leaderDmitry (until he dies) for 5 turns.

\phpaix

\aparag See the explanations hereafter, valid for all turns.

\tour{Turn 2 and afterwards}

\phevnt

\aparag \monarque{Boris Godunov} has a malus of \bonus{+3} to his survival
roll. If he dies, a period of anarchy follows and \RUS has values 3/3/3 as a
monarch. On the next turn, \monarqueRomanov (in fact Fyodor and Michael) is
the new monarch 6/5/6; as this monarch represents the whole family, do not
roll for his survival (it is automatic).

\aparag if \monarque{Boris Godunov} is dead (on this turn or a previous one),
\leaderDmitry also rolls for survival with a \bonus{+3} as sole modifier
during the event.

\aparag As long as the event continues, roll for 3 \REVOLT in \RUS (that occur
only is in a Russian owned province).

\phdipl

\aparag \POL may make a full or limited intervention as ally of the Rebels. It
has a \CB to do so, or a free \CB is \leaderDmitry is alive. This
intervention is not affected by excessive foreign intervention.

\aparag If \POL was involved in this war on the previous turn and \SUE is
making an intervention allied to the loyalists, \POL may generalise the war
with a free \CB in a full war between \SUE and \POL. This does not change the
terms of their respective interventions in the Civil War.

\phpaix

\aparag \REVOLT in provinces that are controlled or occupied by \POL do not
extend and do not count for the conditions of victory of this event.

\aparag If half (round-up) of the Russian national provinces are in \REVOLT ,
\monarque{Boris Godunov} is overthrown and killed with no further
consequences.

% TODO exception to ??

\aparag A side fulfils the military condition of victory if it won a major
victory against the enemy or if it controls all cities in national provinces,
or if the enemy (not their foreign allies) has no \ARMY left.

\aparag The event ends as a victory for the Rebels or the Loyalists under the
following conditions.

\bparag Rebels win if \monarque{Boris Godunov} is dead and they control
\villeMoscou and they fulfil the military condition of victory; or they win if
\monarque{Boris Godunov} is dead and Loyalists surrender willingfully.

\bparag The Loyalists win if all the \REVOLT are eliminated in owned national
provinces and they fulfil the military condition of victory.

\bparag When the victory is obtained, all the \REVOLT and the Rebel armies are
destroyed.

\bparag The intervention of \SUE ends; \RUS has now a free \CB (one use)
against \SUE until the end of the period.

\aparag If the Loyalists win, \leaderDmitry is eliminated.

\bparag If \monarque{Boris Godunov} is alive, he is now legitimate ruler of
Russia. He has now values 6/8/5.  \RUS gains 10\PV.

\bparag If he is not, the new ruler is \monarqueRomanov for 5 turns, with
values 6/5/6. Russian \STAB is increased by 1.

\aparag If the Rebels win, \monarque{Boris Godunov} is eliminated.

\bparag If \leaderDmitry is alive, he becomes tsar \monarqueDimitri with
values 4/7/5 (and the turns left). \RUS loses 3 in \STAB.  If \POL is still
intervening in the war, \RUS is now in mandatory defensive alliance with \POL
during \monarqueDimitri's reign. In addition, \POL gains one province in \RUS
that it currently controls or occupies (its choice).

\bparag If \leaderDmitry is dead, the new ruler is \monarqueRomanov for 5
turns, with values 6/5/6. Russian \STAB is decreased by 2.  If \POL is still
intervening in the war, it gains one province in \RUS that it currently
controls or occupies (its choice).

\bparag In both cases, \POL gains 10 \PV and signs a white peace with \RUS.



\event{pIV:Revolt Cossacks}{IV-17 (2)}{Revolt of the Cossacks}{1}{PB}

\history{1654-1667}
\dure{until the end of the wars caused by the event.}

\condition{If the religious attitude of \POL is Tolerance of Orthodoxy, the
  event does not occur. Mark off and play \RD instead.}

\tour{Turn 1}

\phevnt
\aparag One province of \paysUkraine belonging to \POL (if none, belonging to
\HAB) secedes and create the new minor \paysukraine. The province is
\provincePoltava if available, else, the closest to this one (chosen by the
new protector or controller of \paysUkraine). Units of other countries inside
are immediately redeployed.
\aparag The new minor is a special \VASSAL of its protector. No diplomacy is
allowed on it.
\bparag The protector stops being protector if it declares war to
\paysukraine. The next possible protector in the list becomes the new
protector.
% (Jym) Originally also if does not come to help, but vassal means war to the
% minor is war to the major, so no way not to help
\aparag \paysUkraine never makes a separate peace without its protector and
must be included in the same peace treaty.
\aparag Possible protectors are (in order): \POL (if Orthodox), \RUS, \TUR,
\POL (if not Orthodox). If there are no (more) protectors, \paysUkraine
becomes a normal minor country.

\phdipl
\aparag \POL has a free \CB against \paysUkraine if it loses at least one
province during the formation of that country.
% (JCD): the event does not happen if \POL is orthodox! removing
% \aparag If \POL is Orthodox and does not declare the war, roll for two
% \REVOLT in \POL and ignore the rest of the event.
\aparag If \paysUkraine (as a special Polish \VASSAL) owns a province of
\paysCrimee (a province with a Crimean shield, even blurred), then \POL may
ask for a limited intervention of \paysCrimee in this war.
\bparag This does not change the diplomatic status of \paysCrimee nor its
controller. \paysCrimee is played by its usual controller decided by the usual
rules.
\bparag If \POL wins after an intervention of \paysCrimee, it must give one
province back to it.

\phadm
\aparag If \POL is at war against another \MAJ during the event, \HAB can make
a limited intervention as an ally of \POL.

\effetlong
\aparag \ref{chSpecific:Poland:Polish Ukraine} is no more valid.

\tour{Turn 2 and after}

\phevnt
\aparag If \POL is at war against \paysukraine, \SUE has a free \CB against
\POL.
\bparag If \SUE is at war against \POL, \RUS has a free \CB against \SUE (can
be used in reaction).
\bparag If \RUS uses this \CB and \paysdanemark is either inactive or already
at war with \SUE, then \paysdanemark is put in \EG of \RUS and enters war
against \SUE (if not already at war).

\phpaix
\aparag Normal rules for peace apply, except that allies of \POL cannot annex
provinces of \paysukraine that they didn't own before the event.



\event{pIV:Extension Moghol}{IV-18}{Extension of the Moghol Empire}{2}{PB}

\history{1635-1638 / 1653-1657}

\phevnt
\aparag If the non-European minor country \paysMogol does not exist, it is
created now. Its ruler is now \leader{Grand Moghol} (if period is IV or later,
it replaces \leaderAkbar if he was in play).
\aparag The \paysMogol will try to invade \textbf{3} regions during the turn,
according to \ref{pII:Mughal Expansions}.
\aparag Even if the country has no region after the invasions, it still exists
(and can gain provinces with new events).
\aparag \granderegionBengale and \granderegionKarnatika become rich region,
with 2 resources of each kind shown on the map (instead of 1).



\event{pIV:Wars India}{IV-19}{Wars in India}{2}{PB}

\history{1631-1635 / 1656-1659}

\phevnt
\aparag If it was still existing, minor country \paysVijayanagar is destroyed
(by internal fights).  Every \TP (not \COL) that is in the minor country
\paysVijayanagar at the time of its disappearance will face an attack by
Natives that are activated against every country this turn.
\aparag If \paysVijayanagar had already been destroyed, choose randomly 2 \TP
and/or \COL in \continent{India} that will be attacked by the Natives in the
region, due to internal strife in India.
\aparag \granderegionKarnatika has from now on 2 \RES{Spices} and 2
\RES{Products of Orient} available instead on 1 (if not yet done).
\aparag If the \paysMogol exist, they invade one province with a modifier of
\bonus{-2}, the next in the list according to \ref{pII:Mughal Expansions}.



\event{pIV:Revolt Singala}{IV-20}{Revolts in
  \granderegionSingala/\granderegionFormose}{2}{PB}

\history{1630}

\condition{If there is no \TP/\COL in \granderegionSingala nor
  \granderegionFormose, do not mark off and re-roll.}

\phevnt
\aparag Choose randomly the province of the revolt between
\granderegionSingala or \granderegionFormose if both contain a \TP/\COL. If
not, the chosen province is the one containing the \TP/\COL.
\aparag Place a \REVOLT \facemoins in the chosen region. This \REVOLT is not
connected to the Natives but military forces sent there to suppress it may
have to confront the Natives if they react.



\event{pIV:China Colonial Attitude}{IV-21}{\paysChine Colonial
  Attitude}{1}{PB}

\history{1557 / 1637}

\condition{This event is the same as \ref{pIII:China Colonial
    Attitude}. Exception: if \xnameref{pIII:CCA:Closure China} is already
  effective, apply \xnameref{pIV:CCA:Vassalisation Korea} instead.}


\subevent[pIV:CCA:Vassalisation Korea]{Vassalisation of Korea}

\phevnt
\aparag Two Chinese armies and the natives of \granderegionCorea attack any
\TP/\COL that are in the area (even Japanese \TP).
\aparag \granderegionCorea is now part of \paysChina.

\phpaix
\aparag If a \TP has survived, \pays{Chine} concedes a new \dipAT to the owner
of the \TP, if it didn't have any. The owner still has to pay as for usual
\dipAT with \paysChine.



\event{pIV:Japan Colonial Attitude}{IV-22}{\paysJapon Colonial
  Attitude}{1}{PB}

\history[Tokugawa's Commercial Restrictions in history]{1638}

\condition{}
\aparag If \paysJapon has no \TP, use \xnameref{pIV:JCA:Closure Japan}.
\aparag If \paysJapon has a \TP on the map (in \granderegionCorea), use
\xnameref{pIV:JCA:Japan Commercial Dynamism}.


\subevent[pIV:JCA:Closure Japan]{Tokugawa's Commercial Restrictions}

\phevnt
\aparag One country having a \TP in \paysJapon may sign immediately a Treaty
with \paysJapon, and so gains \dipAT. If more than one country have a \TP in
\paysJapon, all owners (except minor powers) make a secret bidding (minimum
bid is 50 \ducats).  The largest bidder wins and gains the \dipAT; all the
bids are lost and all other \TP are removed from \paysJapon.
\aparag When the \dipAT is accepted, only one \TP of the country is kept in
\paysJapon; excess \TP are destroyed. If refused, \paysJapon declares an
Overseas War against the power.
\aparag From now on, \dipAT allow one country to keep only one \TP in
\paysJapon (and not one per region). The remaining \TP can be upgraded, and it
causes no reaction by \paysJapon.
\aparag The basic forces and reinforcements of \paysJapon are now its mainland
army only (no overseas garrisons or fleets).

\effetlong
\aparag From now on, no new \TP counter can be placed in any area belonging to
\paysJapon by means of administrative actions.
\aparag No regular diplomacy is permitted on \paysJapon.  The Activation level
of \paysJapon becomes 11.


\subevent[pIV:JCA:Japan Commercial Dynamism]{Commercial dynamism of
  \paysJapon}

\phevnt
\aparag \paysJapon gains a \TP with level 6 in \provinceSeoul,
\provincePyongyang and with level 3 in \granderegionFormose, if those
provinces do not contain foreign \TP. \paysJapon has a \FTI of 2, raised to 3
from period V on.
\aparag If there are \TP in any of those provinces, \paysJapon declares an
Overseas War against all the country having those. This war may not be ended
by an automatic white peace.

\phadm
\aparag Basic forces of \paysJapon are increased to 2 \ARMY\faceplus in
\paysJapon, plus 1 \ARMY\faceplus (in \granderegionCorea at start), 2 \LD and
1 \FLEET\facemoins overseas.
\aparag Basic reinforcements are increased to one \ARMY\faceplus in mainland,
and 1 \ARMY\facemoins, 1 \ND for the garrisons.
\aparag If \paysJapon has a \TP counter, it gains 1 level, up to level 6 in
\provinceSeoul and \provincePyongyang, and level 3 in
\granderegionFormose. Choose one randomly for this increase if there are
several \TP.  These \TP exploit the resources in the region and are counted as
normal exploitation for monopolies and evolution of prices.

\phmil
\aparag Japanese forces outside \granderegionJapan do not activate the Natives
and an attack in regions with Japanese \TP may be aimed at the Japanese only
and so does not activate the Natives of the region. As soon as the \TP is no
more Japanese or destroyed, normal activation rules for Natives apply.

\phpaix
\aparag If \paysJapon does not lose the war and there is no \TP in any of the
3 provinces claimed, it places a \TP in there of level 1.

% Jym, 05/2013, Pierre's notes (2007).
\event{pIV:Deluge}{IV-y}{Swedish Deluge}{1}{PBNotEvenWritten}

\history{1648 (Khmelnytsky Uprising)-1667 (Truce  of
  Andrusovo)}[Russo-Swedo-Polish wars, Second Northern war]
\dure{2 turns.}

If \POL is at war, fortresses in \paysLithuanie let enemy supply go through
their province.

Should appear either during IV-17(2), or as IV-10(2).


% Jym, 05/2013, Pierre's notes (2007).
\event{pIV:Koprulu}{IV-z}{\ministreKoprulu}{1}{RistoMod}

Same as~\ref{pV:Koprulu}. Should appear late in the table only. (Jym):
Possibly as IV-17(3) or IV-11(2).

\vfill \pagebreak

\input{engEvntTYW.tex}

\vfill \pagebreak



\event{pIV:Polish Civil War}{IV-B}{Civil War in Poland}{1}{PB}

\history{\textit{alternative history}}
\dure{Until the end of the war}

\phevnt
\aparag Can only happen once, either as explained in \ref{pIV:Liberum Veto} or
in \ref{pV:Saxon King Poland}.
\aparag \POL is now in civil war. One side, called ``Absolutists'' remain
loyal to the King and try to impose Absolutism in \POL while the other side,
called ``Rebels'' is lead by the great nobles of the kingdom trying to keep
the Republic and the elective monarchy.
\bparag The player plays the Absolutists.
\aparag If they have a province bordering \POL, the following countries can
enter a full war against any of the side: \RUS, \SUE, \HAB, \PRU.
\bparag They have a free \CB this turn against both sides of the civil war.
\bparag Other countries can only make a foreign intervention as per normal
rules.
\aparag \textbf{Economic and Political crisis}: The \RT of \POL is diminished
by half and loses at least 50\ducats. \POL loses 2 \STAB.
% \begin{oldcompta}
%   \bparag Do not take into account the minimal loss of 50\ducats.
% \end{oldcompta}
\aparag The Rebels control the following provinces:
\bparag \provinceMalopolska, \provinceLietuva ;
\bparag one other province randomly chosen in \paysmajeurPologne;
\bparag two other provinces randomly chosen in \paysmajeurLithuanie.
\bparag The 5 provinces must be different and all possessed by \POL at the
beginning of the event.
% (JCD): Controlling cities is the same as controlling provinces! removing
% \bparag The Rebels control the cities in these 5 provinces.
\aparag Roll for two \REVOLT in \POL. There are \facemoins and do not control
the cities.
\aparag If \ref{pIV:Times of Troubles} already happened but not
\ref{pIV:Revolt Cossacks} and the religious attitude of \POL is not Tolerance
of the Orthodoxy, \nameref{pIV:Revolt Cossacks} happens immediately.
\aparag The Rebels side is played by the first country at war against the
Absolutists in the following list: \RUS, \SUE, \HAB, \TUR, \HOL, \ANG, \FRA,
\PRU.
\bparag If none is at war against the Absolutists, then the Rebels are played
by the first country in the same list which is not at war as an ally of the
Absolutists.

\phadm
\aparag Lands forces of \POL equal to the basic forces for the period
(excluding Ukraine) become Rebels.
\bparag If \POL does not have enough troops raised, an immediate levy happens,
paid for by the treasure of \POL (even if this causes a bankruptcy).
\aparag The basic upkeep for the Absolutists is the one of \paysPologne only
(\ARMY\faceplus).
\bparag The player may use the counters of \paysPologne and two \ARMY (they
can be taken from any unused country, and are similar to any other Polish
Army)
\bparag Absolutists receive normal income from the provinces they control.
\bparag Absolutists troops in rebel provinces (at the beginning of the war)
are retreated normally.
\bparag Fleet stay loyal to the Absolutists.
\bparag The king of \POL must be used as a general of the Absolutists, except
if he is \monarque{August II}.
\aparag The Rebels side uses the counters of \paysLithuanie as well as two
revolts \ARMY.
\bparag He does not get reinforcement at the first turn of the war.
\bparag At the first turn of the war, the Rebels forces can be freely
redeployed in the controlled provinces.
\bparag If a named general (other than \leaderPatkul when \monarque{August II}
is king) is in play, he takes side for the Rebels. Otherwise, the Rebels are
lead by a random mercenary general and get an extra random general.
\aparag \REVOLT in \POL are friendly to the Rebels.
\bparag A rebel general can lead a \REVOLT . A \REVOLT \facemoins count as
2\LD for hierarchy rules.
\aparag Starting with the second turn of the war, Rebels get reinforcement
either in offensive or defensive attitude based on the income of the province
they control (control the city with no absolutist army in the province).
\aparag If \paysukraine is not in revolt or independent due to \ref{pIV:Revolt
  Cossacks}, the Ukrainian army can be used by the Absolutists (but without
the basic upkeep for it).
\aparag If the king is member of the dynasty of \paysSaxe, he can use the
forces of the minor as per the rules of \ref{pV:Saxon King Poland}.
\bparag In that case, \paysSaxe is at war against the Rebels and their allies
can freely cross the \HRE and wage war in \paysSaxe.

\phmil
\aparag Absolutists and Rebels get supply from the cities they control.
\bparag They can cross enemy provinces without besieging the city.
\bparag This is only true for polish forces. Not for the foreign allies.
\bparag The Absolutists cannot cross freely the provinces with a \REVOLT .

\phpaix
\aparag Victory in the civil war occurs as soon as one side gets two out of
the following three conditions:
\bparag controlling the capital (controlling \provinceMalopolska and, if
\villeVarsovie has been made capital, \provinceMazowia) ;
\bparag controlling the country (military control of at least 60\% of the
provinces, that is controlling the city without enemy presence ; provinces
with a \REVOLT and the city still controlled by the Absolutists count for
nobody) ;
\bparag military victory (having one more major victory than the other side
this turn, or the other side as no more \ARMY in play).
\aparag The war lasts as long as no side wins.
\aparag Wars with foreign countries can be ended by separate peaces.
\bparag If the Absolutists are not fully at war against another major country,
\POL does not lose \STAB due to the war (but does so due to \REVOLT ).
\bparag A (foreign) peace in the civil war is also a peace with \POL (if
another war was occurring), or a separate peace with loss of 2 \STAB for
breaking the alliance with the side of the civil war the foreign country was
allied to.
\aparag[Absolutists victory]
\bparag The effect of \ref{chSpecific:Poland:Liberum Veto} are cancelled.
\bparag Events \xref{pVI:Great Northern War}, \xref{pVII:Bar Confederation},
\xref{pVII:First Partition Poland}, \xref{pVII:Second Partition Poland} and
\xref{pVII:National Revival of Poland} are modified.
\bparag Any country fully allied with the Absolutists who accept the peace
annexes a province of \POL (\POL choose which).
\bparag The Rebels armies are eliminated.
\bparag The \REVOLT stay in place.
\aparag[Rebels victory]
% (Jym): added condition on election
\bparag A dynastic crisis occurs and a new king is elected (this is a change
of polish dynasty), a general cannot be elected king unless he took the side
of the Rebels.
\bparag A Polish provinces is given to each \MAJ who was fully at war against
the Absolutists (choice is made by the \MAJ receiving the province, in order
of initiative).
\bparag The \STAB of \POL immediately becomes -1.
\bparag The \REVOLT and the Absolutists armies are removed.

\stopevents

% Local Variables:
% fill-column: 78
% coding: utf-8-unix
% mode-require-final-newline: t
% mode: flyspell
% ispell-local-dictionary: "british"
% End:

% LocalWords: pIV TYW HRE offensive reroll minister Olivares Safavids pIII pV
% LocalWords: Liberum Oxenstierna Torstensson Moghol Singala CCA PBNew JCD de
% LocalWords: Vassalisation Bethlén Mansfeld Westphalie Ausgsburg Schmalkaldic
% LocalWords: malus Risto Espagne HOL Ormus POR Jym reannexed Reannexation
% LocalWords: RistoMod Angleterre ECW Montrose Duche Prusse Pommern Ost JCA
% LocalWords: Torstensson's Fyodor defensive Formose pII Alaouite pVI pVII
% LocalWords: PBNotEvenWritten willingfully
