% -*- mode: LaTeX; -*-

\definechapterbackground{Expenses}{expenses}
\chapter{Expenses}\label{chapter:Expenses}
\begin{designnote}[Expenses] This chapter describes mainly all the logistics
  (maintenance and purchase of new forces), country management and development
  operations. All of these form the administrative phase, that may be cut into
  three main sections: administrative operations, automatic actions and
  logistics. Administration and Logistic are conducted independently and
  simultaneously by all players while automatic actions resolve the conflicts
  that arise from the new state (incompatible placements of new colonies,
  monopolies, etc.) Most of the parameters of administrative operations were
  explained in~\ref{chapter:ThePowers}, and serve mostly
  in~\ref{chapter:Incomes} and this chapter. The nature of military troops is
  explained here before the logistics details.
\end{designnote}

\begin{playtip}
  The easier way to play the administrative phase is to have all players
  simultaneously compute their income and plan their expenses (both
  administrative and logistic). Once all expenses are planned, players can
  resolve them two by two (one player resolves his action with the other
  watching the die rolls).

  Often, one will find out that planning logistic ahead (prevision for
  campaigns cost, then buying troops and fortresses) has do be done before
  planning administration as military can be the priority expense and
  administration only gets the ``leftover'' money. Of course, this must be
  based only on an estimate of the income, since the exact amount will only be
  known at the end of the turn, after the Exchequer test\ldots

  The most efficient way to plan and resolve actions is to write everything on
  the record sheet: nature of the actions, investment, column and bonus. Thus,
  resolving the action is very fast: simply announce column and bonus and roll
  one die with the other player checking the result in the table. Writing down
  the result of your actions (success, failure, or amount of progression
  gained) will allow to quickly resolve all the actions in a row. Then, you
  can ``release'' the other player and actually implement the results obtained
  (which can require finding the right counter and similar things).
\end{playtip}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\aparag[Sequence.]
\AdminDetails

\input{engExpAdmin}

\input{engExpLogistic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Logistics in French}
% {\bf Dans V.4}

% {\bf E1 placement rotw}

% Les E et C peuvent être placés en début de tour en métropole, ou dans une
% COL/TP établie, ou dans une région découverte où se trouvait un chef du même
% type à  la fin du tour précédent.  Les Gouv et G(rotw) doivent être mis sur
% des COL/TP déjà  établis.  Les missionnaires restent en place sans
% redéploiement, et apparaissent toujours en métropole.

% {\bf E2 placement des non nommés}

% Les chefs ? sont tirés au début de chaque tour pour combler si en dessous de
% la limite et tous les chefs sont placés en même temps en respectant la
% hiérarchie (seuls les chefs pris dans un siège n'ayant pas de port sans
% blocus restent nécessairement sur place).

% Si pendant les rounds, on tombe à  nouveau en-dessous des limites, un chef ?
% de la bonne catégorie est tiré pour remplacer le dernier de ce type tué ou
% blessé au round précédent. Il doit être mis au même endroit pour les A et G;
% pour les E et C, on les met en métropole ou sur un TP/COL; ensuite seulement
% on doit règler les pb de hiérarchie comme on peut. Quand un blessé revient,
% le chef non nommé de sa catégorie de rang le plus bas est enlevé, et le chef
% revenant doit prendre commandement d'une pile en respectant la hiérarchie.

% {\bf E3 chefs gouverneurs / changement de C}

% Certains généraux ou C sont en fait des gouverneurs.

% Ils agissent comme des C mais entrent dans une limite de chefs
% spécifiques. Ils obéissent de plus aux règles suivantes : - ils doivent être
% placés sur un TP/COL du joueur ou qu'il occupe militairement en début de
% tour, ou en métropole. Si en métropole, ils doivent débarquer dans rotw sur
% TP/COL du joueur (ou contrôlé militairement).  - ils donnent un bonus de
% +man aux implantations dans leur province, et de +1 dans les autres
% provinces de la grande région.  - ils sont limités aux mouvements dans la
% grande région où ils sont placés en début de tour (ou celle où ils
% débarquent) et les grandes régions adjacentes.

% Local Variables:
% fill-column: 78
% coding: utf-8-unix
% mode-require-final-newline: t
% mode: flyspell
% ispell-local-dictionary: "british"
% End:

% LocalWords: Carrack Tercios Galleass Boyars
