% -*- mode: LaTeX; -*-

\definechapterbackground{Playing the game}{chess}
\chapter{Playing the game}\label{chapter:Playing}

\begin{todo}
  Technical details and advice: how to sort the counters, build the player
  aids, use the generic record sheets (TF, exotic resources, \ldots) and
  generally how to hold a game session.
\end{todo}

\section{Building the Game}

\section{Playing the Game}

\section{Teaching the Game}
This Section explains how to hold a Tutorial session to teach the game (on 1
or 2 days). It's a more a checklist of things I say when teaching the game
than anything else. Hence, it is somewhat terse and is intended for people who
already know the game and want to teach it (not for people who don't know the
game and want to learn it, although it can certainly be used as a large
scenario). Be sure to read this Section fully before the Tutorial (so you know
what to say and not say) and use it as a checklist during the session. Note
that due to all explanations and sequential solving of actions (everybody want
to see what's happening to understand a bit), Turn 1 might well last for 4 to
6 hours even without any real Diplomacy\ldots

Most of the tutorial happens on the European map. Thus it is very well
possible to play with only one map (takes less space). In this case, it is
advised to not play \POR, and abstract the \ROTW for \HIS (playing it only on
the minimap). You'll still need a \STAB track and the first line of the
technology track (draw them on some paper). Prices of exotic resources won't
change, but that's not an issue (that is, they should change but you won't be
able to track that so you'll play with fixed prices, which is good enough for
teaching the game). You'll also need a European Diplomatic track.

It is possible to play with more or less than 9 players. If playing with more
than 9, make teams. If playing with less, the focus should be around the
Mediterranean: \TUR, \VEN, \FRA, \HIS. Next, add \POL and \RUS (if 2 more
players). \POR has a mostly solo play but is interesting to teach some \ROTW
aspects (together with \HIS). \ANG and \DAN can be left out without
troubles. If using this as a scenario for learning the game, focus on Italy
playing only \FRA, \VEN and \HIS (and not the \ROTW part).

\subsection{Before the Tutorial}
\aparag Make sure to print everything needed\ldots
\bparag As far as possible, get the counters glued and cut in advance.
\bparag The \TradeFLEET Incomes (on the Colonial record sheets) need to be
manually filled (they can't be pre-filled because they're likely to
change). Do it before the Tutorial.
\bparag Same goes with the global record sheets (\TradeFLEET and Exotic
Resources).
\bparag Set up the game before starting to explain, if needed, do the setup
before the rest of the players arrive\ldots There are quite few counters in
play on Turn 1.

\aparag Initial placement
\bparag Make sure that \FRA has enough troops in \provinceProvence to reach
2\ARMY\faceplus there after recruiting (that's 12\MP to \villeNaples).
\bparag Do not place \HIS troops in \provinceCatalogne (to avoid speaking
about troops in revolted provinces) but still close to it (so it can move
there on first round).

\subsubsection{Guideline}
The idea is to introduce stuff only when needed. The first
approaches (explaining material, going through the turn phases) should stay at
a very high level (\emph{e.g.} say that provinces' colour is terrain but do
not try to explain the differences between each kind of terrain as it would be
pointless at that point). Later, when actually playing through Turn 1 and Turn
2, detailed explanations become needed and start to make sense.

Of course, if players are curious and ask questions, answer them. But try to
avoid getting into details too early as you'll likely lose them. That is a
perfect answer is ``this is resolved by die roll and reading a table, we'll
see that in details later when we actually do it.''

As the teacher, do not be shy to take some decisions instead of the players
(\emph{e.g.} undermining or assault?) as it can be difficult for newcomers to
fully understand the implications while the choice looks obvious for an
experienced player. Taking decision in their place will avoid too much
Analysis Paralysis. You may even take ``bad'' decisions at some point to show
why it is bad (``assaulting a fortress of level 3 in \TMED? Sure, let's do it
and figure out why that is silly.'') Be sure to warn the players in advance
(``This is a bad choice, let's do it to see what happen''); and don't be shy
to cheat on some rolls (``suppose you had rolled 8 instead, \ldots'') or
rollback some choices afterwards (``OK, that was silly, let's go back and take
another option''). Of course, warn the players when starting that you might
sometimes do that to show stuff\ldots

One the other hand, do not take all decisions immediately and leave the
players with a feeling of being here only to push counters. As much as
possible, try to offer and explain two alternatives (usually, low risk low
reward or high risk high reward) and give them a bit of time to make a
choice. Since this is a tutorial, doing ``stupid'' things is not stupid at
all\ldots

The goal of the Tutorial is only to learn the game. There is no winner in
it. Everything that happens during the Tutorial will be erased if you start a
real game. Thus, it's better to see a variety of situations rather than
optimise your choices.

\subsection{Initial Explanations}
\subsubsection{Start of the Tutorial}
\aparag Global organisation of tutorial session:
\bparag Quick go through material, aids and rules
\bparag First turn is scripted and stuff will be detailed when playing it.
\bparag Some things done during T1 may be stupid but serve pedagogical
purpose\ldots
\bparag Many special rules are skipped for tutorial.
\bparag Turn 2 is not scripted, but the teacher will make some complex
decisions instead of the players (in order to do so instantaneously).

\subsubsection{Explain material}
\aparag Map
\bparag Provinces: income, fortresses of level 1 or 2, terrain.
\bparag \ROTW: Areas with provinces sharing stuff (coloured boxes and
borders).
\bparag Sea zones and \CTZ/\STZ

\aparag \STAB
\bparag Several tables on \ROTW map, notably \STAB.
\bparag \STAB is super important for incomes and peace.
\bparag \STAB is \textbf{super important}.

\aparag Counters
\bparag \Faceplus/\facemoins concept. Mostly for economical counters (\COL,
\TP, \TradeFLEET), but also for \ARMY and \FLEET.
\bparag Show leaders, do not explain values.
\bparag Skip the rest.

\aparag Specific player aid
\bparag Country player aid: period limit table, turn limit table. Only the
first line of these (period \period{I}) is used for tutorial
\bparag Troops and navies cost. Only the first two lines are used for tutorial.
\bparag Recall of Specific rules: mostly ignored for tutorial.
\bparag Do not explain what each column of the table means, just that these
are related to how many counters you may have in play, or how much your army
cost, but that's it.

\aparag Monarch sheet
\bparag Monarch values, initiative
\bparag Period and turn limit reminders
\bparag Actions planners (pre-filled for T1, do not explain yet what they
mean).

\aparag ERS
\bparag 3 ERS on 2 sheets: RT, income and expenses, loans

\aparag Colonial sheet
\bparag Colonies, \TP, \TradeFLEET are recorded here.

\aparag Common player aids
\bparag sorted in turn order
\bparag many, many tables, skip them now
\bparag go to page 8 with game sequence

\subsection{Going through the game turn}
\aparag Using the Game sequence, go through game turn and explain a bit what
happens there.
\bparag Only explain the concepts, not the precise rules and resolution.
\bparag \emph{e.g.} for diplomatic actions, say that you can spend money to
influence minor countries and this is resolved by dice, but do not explain all
the modifiers and stuff at this point. As a rule of thumb, the explanations
here should stay high level enough that actually looking at any of the tables
is useless.

\aparag Event phase.
\bparag Check if monarchs die
\bparag Random economical event (1 per country)
\bparag Random political events. They do give the game tempo.

\aparag Diplomatic phase
\bparag Alliances between \MAJ.
\bparag Declaration of war (cost \STAB, and \STAB is super important).
\bparag Diplomatic actions on minors.
\bparag Discussions can take a lot of time before players reach an agreement,
Diplomacy is an important part of the game\ldots

\aparag Income phase
\bparag 4 sources of income: land, industry, trade and \ROTW.
\bparag Just explain what each category corresponds to globally, no details.
\bparag Gross income is not real income, there will be an Exchequer test.

\aparag Administrative phase
\bparag Economical development of your country.
\bparag Buying new troops.

\aparag Military phase.
\bparag Rounds, impulses (in initiative order).
\bparag Impulse: move all stacks, resolve all battles.
\bparag Movement may cause attrition.
\bparag Interceptions and counter-interceptions are possible.
\bparag Sieges are resolved after all impulses.
\bparag This is a long phase (several rounds) and a complex one (many
decisions to take). After all, the game is largely a wargame\ldots

\aparag Redeployment phase.
\bparag Mostly cleanup of Military phase.

\aparag Peace and Budget phase.
\bparag Signing peaces (\STAB is important as it decides level of peace!)
\bparag Exchequer test.
\bparag Explain the concept of cutting the income in three.
\bparag Explain the idea of the Prestige income that must be spent for VPs.
\bparag Explain that part of income has to be borrowed (loan) and you may go
bankrupt if you don't want to refund it.
\bparag The economical system works better with loans.
\bparag Exchequer test is complicated and is a complex and meaningful decision
to take each turn.
\bparag \STAB is important for Exchequer test!
\bparag \STAB improvement. Paying money is the best way to get \STAB.

\aparag Interphase
\bparag Cleanup for the turn.

\subsection{Turn 1 -- Event and Diplomacy}
\aparag This is tutorial
\bparag Some Segments are skipped, will be added later.
\bparag All actions and decisions are already scripted.

\aparag Whenever you need to resolve an action or compute a value, go to the
correct table in the aid and explain the process in details.
\bparag But there is no need to read all the modifiers of complex tables
(\emph{e.g.}, when rolling for land attrition read the modifiers headers and
skip the parts that only deal with naval attrition).
\bparag Be sure that the players understand each computation of modifiers and
reading of tables. Go slowly the first time. Let them do it after that (always
checking that they do it correctly).

\subsubsection{Event phase}
\aparag Only political events, and they are pre-rolled.

\aparag Do not read the full events, just give an overview of the decisions.
\bparag Think about ticking off boxes in the event table that can produce
these results\ldots

\aparag \ref{pI:War Italy Napoli}
\bparag \FRA has claim on \pays{naples} (aka Napoli) and attacks it. \HIS has
family there and is not happy.

\aparag \ref{pI:Habsburg Dynasty}
\bparag Dynastic actions are a specific thing for \HIS.
\bparag \HIS does \ref{pI:Habsburg Alliance}. Check box on player's
aid. Explain that player's aid hold spaces to remember some country specific
stuff.
\bparag \HIS is now ally with \paysHabsbourg (aka Austria).

\aparag \ref{pI:War Scotland}
\bparag \paysEcosse attacks \ANG.
\bparag If \ANG is not played, still do this event but explain that it will
have no effect\ldots

\aparag Revolt
\bparag \REVOLT\facemoins in \provinceCatalogne.
\bparag Do not speak of the Diplomatic event that normally comes with it. Do
not show the revolt tables, just say that it has been randomly decided.

\subsubsection{Diplomatic phase (Majors)}
\aparag Wars due to event
\bparag \FRA attacks \paysNaples. \HIS join in Limited
Intervention. \paysSavoie joins \FRA because of ``specific rules of the
event'' (do not make the checks for the other Italian countries, do not speak
of Venetian possibilities).
\bparag Lower \STAB accordingly. Recall that \STAB is super important.

\aparag Discussions: Explain that normally players can discuss privately and
make agreements but that is already decided. Plans for each country for the
turn:
\bparag \ANG vs \paysEcosse.
\bparag \FRA in Italy, needs to take \villeNaples for \VPs due to event.
\bparag \HIS crushes \REVOLT, plus maybe tries to help \paysNaples. Plus goes
to \continentAmerica.
\bparag \POL and \RUS ally and try to destroy \paysDon.
\bparag \POR explores and tries to go to \continentIndia.
\bparag \TUR attacks \paysGeorgie and tries to moves in the \regionBalkans
(it's free for all).
\bparag \VEN tries to take \villeRagusa (it gives a commercial bonus, do not
try to explain it).

\aparag Announcements.
\bparag \RUS and \POL sign an offensive alliance.
\bparag Explain Defensive and Offensive alliances.

\aparag Declarations of War (1).
\bparag \RUS attacks \paysDon, \POL joins the war.
\bparag Lower \STAB (\RUS has no \CB, \POL has due to alliance).
\bparag Explain the deal: the first annexed province goes to \RUS, the second
to \paysUkraine, the third to \RUS. That's why \RUS is declaring the war
first (and looses more \STAB).

\aparag Declaration of War (2).
\bparag \TUR attacks \paysGeorgie.
\bparag Explain that \TUR has \CB due to religion. Lower \STAB.

\subsubsection{Diplomatic phase (Minors)}
\aparag Explain the correct way to do it:
\bparag Secret planning of action and writing (target + investment + final
modifier).
\bparag Rolling dice, writing results (for all actions).
\bparag Changing game state (moving markers).

\aparag For tutorial, all actions are written. Be sure to resolve them in the
order indicated here.
\bparag Before resolving the first action, explain the principle: 1d10+mod vs
2d10, gaining progression points.
\bparag Remember: this is a game with \textbf{many} die roll. One diplomatic
action may look random, the large scale course of diplomacy on minors isn't.
\bparag Before resolving each action, make the player compute the modifier for
it. Take the time it need to find the correct modifier.
\bparag After rolling the dice for each action, make the player write down the
result.
\bparag After rolling (and noting) for the first action (and if it's a
success\ldots) grab the diplomatic counter and explain how to make the
progression. Don't hesitate to cheat on the first action so that it becomes a
success and you can explain (``let's say the minor only rolled 3 and 6,
\ldots'')
\bparag If reaching \SUB or \RM, explain the choice and let the player
choose (they may stop at a lower box or let the minor stay Neutral).
\bparag Let the players try to do the progression themselves, but check that
it is correct.

\aparag \POL on \paysBrandebourg.
\bparag Explain the principle. Compute the modifier. Check that it is the same
as written.
\bparag Talk about various investments and say that it is usually better to
keep low investment for most actions.
\bparag Roll dices. Write result.
\bparag If success, grab the counter, explain that the values are progressions
points for the various levels and do the progression.

\aparag \POR on \paysMaroc and \VEN on \paysPerse
\bparag Show how the religious bonus/malus is important (compare with previous
action).

\aparag \RUS and \TUR on \paysKazan.
\bparag Explain what happens when two countries attempt diplomacy on the same
minor and resolve it.
\bparag Because of that you need to secretly plan everything, then announce
everything (and find conflicts), then start resolving.

\aparag \TUR on \paysAlgerie and \paysDamas.
\bparag At this point, players should be able to resolve actions on their
own. Let them do the steps while simply checking that everything is done
correctly.
\bparag There's a \bonus{-3} specific malus for \TUR on Barbary coast
countries at the start of the game.

\aparag \HIS on \paysPapaute, \paysSuisse and \paysLorraine.
\bparag Let \HIS do its stuff, just tell them not to do the last written
'action'.
\bparag \paysSuisse has no religion (in game).

\aparag \FRA on \paysGenes.
\bparag This is an attack on \HIS minor and that \HIS can (and will)
react. Because this is a reaction, \HIS cannot gain levels even if winning.
\bparag Resolve action.

\aparag \FRA on \paysToscane.
\bparag There is special a \bonus{+2} due to the event.

\aparag Don't forget to update \EcoRS A according to results. Make sure every
player has correctly filled its \lignebudget{RT after Diplomacy}.
\bparag It is unlikely that someone gains a new \VASSAL but if it happens
remember to handle that when computing Income.

\subsection{Turn 1 -- Incomes and Expenses}
\subsubsection{Income phase}
\aparag Go through \EcoRS B Section by Section (and line by line). After
explaining each, give the players enough time to compute if themselves (except
maybe for Provinces Income which can be difficult\ldots)

\aparag[Land Income]
\bparag \lignebudgetlong{Provinces income}: can be difficult to compute but
does not changes often (so just copy old value). Do not forget \provinceAcores
and \province{Islas Canarias}.
\bparag \lignebudgetlong{Vassal provinces income}: Do not forget to update if
there's new \VASSAL after Diplomacy.
\bparag \lignebudgetlong{Occupation, Pillages, Revolts}: subtracting these
allow to not change \lignebudget{Provinces income} very often.

\aparag[Industrial Income]
% \bparag \lignebudgetlong{Manufactures}
\bparag \lignebudgetlong{European mines}: \POR gets 40\ducats for
\construction{Elmina}.

\aparag[Trade Income.] Many of these computation are not very easy. Take your
time to have everybody managing to find the good value.
\bparag \lignebudgetlong{Domestic trade income}: explain \DTI (and \FTI),
special power of Cereals \MNU.
\bparag \lignebudgetlong{Foreign trade income}: special power of Wine and
Clothes \MNU.
\bparag \lignebudgetlong{STZ+CTZ level income}: difference between \STZ and
\CTZ.
\bparag \lignebudgetlong{STZ+CTZ monopoly income}: what is a Monopoly in
\STZ/\CTZ.
\bparag \lignebudgetlong{Partial/Total monopolies (trade)}: recorded for \VPs
purpose.
\bparag \lignebudgetlong{Trade centres income}: \CCs{Grand Orient} has variable
value (do not explain computation, just give value) and is not related to
\TradeFLEET. Explain that it's shared between \paysEgypte and \paysDamas and
in turn they sometimes give it to \VEN and \TUR.

\aparag[\ROTW income]
\bparag \lignebudgetlong{Colonies}: Explain a bit the 3 values of an \Area
\bparag \lignebudgetlong{Exotic resources}: in the \ROTW, they are shared for
an \Area. In Europe, Fish and Salt \MNU produce some. Special Salt \MNU of
\VEN.

\aparag[Gross income] Immediately copy it back in \lignebudgetlong{Gross
  income A}.

\subsubsection{Administrative actions}
\aparag Again, explain the correct way to do it:
\bparag Compute and write everything (action, investment, column, modifier).
\bparag Roll for all, simply writing result.
\bparag Once all rolls are done, go and change state.

\aparag And be sure to resolve it that way\ldots
\bparag Have one country make all its rolls, noting results.
\bparag Only them show how to implement the results.

\aparag Explain the general principle of administrative actions.
\bparag One table to resolve them all. Column and DRM.
\bparag Ways to compute the column and DRM vary widely for each action.

\aparag As for Diplomacy, go slowly, explain how to compute (that is, how to
read the table). Be sure that every player manages to compute its own columns
and DRMs and find the correct values.
\bparag Skip special cases (critical failures in \ROTW actions).
\bparag It's not an easy table to read the first few times, and it is full of
abbreviations. So give the players the time and help they need to figure it
out.

\aparag \POR is a good country to start with as it does a bit of everything.
\bparag Explain the Special \FTI.

\aparag Most actions are done with small investment.
\bparag This maximise long term chances of success.

\aparag Some actions (Technology) are done with medium or strong investment.
\bparag It is actions where there can be an huge immediate gain (new
technology) and thus a short term boost is important.
\bparag Show how a small investment would mean way less chances of immediate
success.
\bparag Sometimes, it's just worth paying a lot.

\aparag Don't forget about conversion cost for new technology, but don't
forget either that it has to be paid for the newly recruited troops.

\aparag When implementing the result of the action, make sure to immediately
change the \EcoRS \textbf{B} for next turn.
\bparag Insist that this is one of the most important trick in keeping the
Income phase manageable.
\bparag Empty boxes mean ``same as previous turn'' and are easy to copy. As
soon as something changing Income happens, it should be recorded in \EcoRS
\textbf{B} so it's not forgotten.

\subsubsection{Logistic}
\aparag Talk about recruitment limits and make sure that everybody can
correctly compute the price of their troops.

\aparag \ANG recruits 2\ND and 3\LD.

\aparag \FRA recruits 1\ARMY\faceplus.
\bparag Be sure that \FRA has 2\ARMY\faceplus in \provinceProvence.

\aparag \HIS does not recruit anything.

\aparag \POL recruits 1\ARMY\faceplus of \paysmajeur{Pologne}.
\bparag Explain briefly that \paysmajeur{Pologne} and \paysmajeur{Lithuanie}
are two countries both part of \POL.

\aparag \POR recruits 2\ND.

\aparag \RUS recruits 1\ARMY\faceplus.
\bparag Talk a bit about recruitment areas, and the fact that \RUS can only
recruit in \villeMoscou.

\aparag \TUR recruits 1\ARMY\faceplus of \Janissaire.
\bparag Talk about the difference between \Janissaire and \Timar.

\aparag \VEN recruits 1\ARMY\facemoins.

\aparag Make sure to compute now conversion costs for new technologies and
write them down.

\aparag Give leaders to countries.
\bparag Talk about leader limits, named leaders and anonymous leaders, types
of leaders (\LeaderG, \LeaderE and \LeaderC are going to be the most important
on turn 1).
\bparag Give each country its leaders.
\bparag Placement of leader should be fairly obvious, but don't hesitate to
help players\ldots
\bparag Mention Hierarchy if needed, but it should be good. Talk also a bit
about \Pashas, but on turn 1 they probably don't need to be used as leaders,
so don't insist on that point.

\subsubsection{Administration of minors}
\aparag Minors at war get reinforcement.
\bparag this is in addition to the \terme{Basic forces}.
\bparag First choose an attitude, then roll die.

\aparag \paysEcosse has to choose \terme{Offensive} (because of the event),
\paysSavoie chooses \terme{Offensive} attitude, all other choose
\terme{Defensive}.

\aparag Placement of minors.
\bparag \paysCosaquesdon should put its troop in \provinceDon (behind the
river) and its fortress in \provinceDon if getting only one (last ditch stand)
or in both \province{Dikoe Pole} and \provinceDonets if getting two (wall to
slow the enemy and troops behind the river). As usual, don't hesitate to do
the placement instead of the players.
\bparag \paysNaples may place its troops either in \provinceCampania to
protect \villeNaples (risky, but it's also a good occasion to show a battle
and maybe explain major battles\ldots) or in \provinceBasilicata (waiting for
Spanish reinforcement before counter attacking, but if \villeNaples falls
early, this is pointless). The first fortress level should go in \villeNaples,
more don't really matter.
\bparag \paysGeorgie should be placed so that it can hide in its fortress
(thus, not 2\LD in \provinceGeorgie if it didn't get a fortress as
reinforcement\ldots) First fortress goes in \provinceGeorgie (\TUR goal is to
take everything, so one very difficult province is better than two mildly
difficult).

\subsection{Turn 1 -- Military}
\aparag There's not going to be a lot of action on turn 1. Mostly sieges.

\aparag Compute Initiative for each alliance (\POL and \RUS are now together).

\aparag Roll to determine first round. For the first turn, ignore bad weather,
extended campaigns and the Wintering Segment.

\subsubsection{Round 1: Impulses}
\aparag The first move should be \FRA going to \villeNaples.
\bparag Explain how to compute \MP, talk about \MP limit. See if the players
can find the various 12\MP moves from \provinceProvence to
\provinceCampania. Some are better than the others since they do not require
to leave troops behind for a siege in \provinceAbruzzo.
\bparag Resolve the Attrition. Attrition table is quite complicated to read,
so go slowly.

\aparag Depending on where the troops of \paysNaples are, there might be a
battle in \provinceCampania or not. If there is one, take the time needed to
resolve it correctly. Note everything in clear view of everybody and so
on. The teacher can act as an external ``secretary'' and write down the result
while the concerned players roll the dice and read the table (same goes for
the first few other battles). Battles tables are also quite difficult to get
used to; reassure the players telling them that it happens often enough that
they'll get it quite quickly.

\aparag The rest of the French troops, as well as the Savoyard, can move
slowly. Especially, make sure that \FRA does not end up paying a 100\ducats
campaign while moving the reinforcements\ldots

\aparag \HIS should go crush its \REVOLT and start exploring the Atlantic.

\aparag \POR explores the African Coast.

\aparag \VEN go besiege \villeRagusa.

\aparag \POL and \RUS attack \paysCosaquesdon. Typically with \RUS in
\province{Dikoe Pole} and \POL in \provinceDonets.

\aparag \TUR attacks \paysGeorgie (troops hide in the fortress) and maybe also
move in the \regionBalkans. But that's more expensive\ldots

\aparag \ANG need to first kill the Scottish troops and then make siege.

\aparag Make sure to check Attrition for these moves. At sea, and for large
stacks on land.

\subsubsection{Round 1: Sieges}
\aparag There should be a lot of sieges going on. Explain the general choice
between undermining and assault, then resolve them one by one. Computing the
Artillery bonus is the tricky part.
\bparag Don't hesitate to make the choice instead of the players to speed up
the game.
\bparag Of course, you can also make an example at some point, \emph{e.g.} if
there are only been Undermining so far, be like ``let's try an Assault just to
see how it goes''.

\subsubsection{Subsequent rounds}
Things can vary wildly depending on what happened previously and the result of
the actions\ldots

\aparag In Italy, it depends on how fast \HIS crushes its revolt and whether
the Sicilian troops are still alive or not.
\bparag It makes little sense to try and recapture \villeNaples if already
taken (situation will be way more favourable for the Spaniards next turn), but
it does make sense to try and lift the siege.
\bparag Similarly for \FRA, the main goal in the war is to take
\villeNaples. Only once it's taken can they start trying fancy stuff (taking
more provinces, killing troops, \ldots)

\aparag In the \regionBalkans, things should move fairly slowly. It is likely
that \VEN does not manage to capture \villeRagusa.

\aparag On the Don, or in Georgia, things will highly depend on the siege
rolls and the length of the turn. The war in Georgia should be fairly obvious
to handle.
\bparag As soon as one province of \paysCosaquesdon is taken, the
corresponding country should move forward to \provinceDon. There will be a
battle here. Since it's likely the first or second battle of the game, go
slowly.

\aparag In the \ROTW, \HIS tries to reach \granderegionCuba (best reward/risk
ratio for early colonisation) and \POR tries both to explore the Eastern
African coast (so it can attack \paysGujerat next turn) and reach
\continent{India}.
\bparag Note that they will get a free repatriation move at the end of the
turn, thus there is no need to go back with the discoveries each round.
\bparag Early exploration is meant to be quite costly in vessels and troops.
\bparag Try to make sure that they both leave a \LeaderC in a good spot for
\COL/\TP action on turn 2 (likely \granderegionCuba for \HIS, for \POR it may
be \province{Natal N} or \province{Karnatika S} which are both good starting
points for further expansion).

\subsection{Turn 1 -- Redeployment}
\aparag Don't loot.

\aparag Spanish and Portuguese fleets probably need return to port with the
discoveries. It might be extremely costly for \POR (crossing the Cape of Good
Hope). It's normal. Don't forget to leave a \LeaderC in good spots.

\aparag Some failed sieges might need redeployment. Likely in \villeRagusa due
to the \regionBalkans rule. Maybe in the other wars.

\aparag In the unlikely case where the \REVOLT is still here, just cheat and
remove it for free\ldots

\subsection{Turn 1 -- Budget}
\subsubsection{Exchequer test}
\aparag This is one of the most difficult decision in the game. The choices
are extremely meaningful and can lead to very different results. And the way
things happen is unusual enough that it is hard to figure out the implications
of each choice.
\bparag At the same time, the tutorial is the time to make bad choices and not
suffer from them.
\bparag It is also the only table out of sequence in the players aids (due to
re-using the Administrative table).

\aparag The mid-term goal is to have a bit less than 100\ducats in Treasure
(50\ducats for ``poor'' countries) and a bit less than 100\ducats in national
loans. That's how the system can be used at best.
\bparag Of course, circumstances can create other short term priorities.
\bparag Also, countries that start high on cash (\emph{e.g.}, \POR) need to
make it last a bit, especially while inflation is low.

\aparag First, explain the roll itself. Let the players compute the modifiers,
make the rolls and fill the lines.
\bparag Explain that each box is split in two so one can write both the
percentage and the actual value. Make sure lines \lignebudget{Regular income}
to \lignebudget{Max. national loan} are correctly filled.
\bparag Skip International loan, they shouldn't be needed.

\aparag Next, explain how the expenses have to be paid:
\bparag \terme{Regular Income} must be used to pay expenses, as much as
possible.
\bparag \terme{Prestige Income} is choice between paying expenses or making
\VPs.
\bparag \terme{National loans} is debt but can be used both for paying and for
refilling the treasure.

\aparag Let the players make choices. There can be bad choices, but for a
Tutorial they will have no real consequences (and players can understand why
they're bad only by doing them).
\bparag Take special care of correctly filling \lignebudget{Remaining
  expenses} to \lignebudget{from N. loan}. They are a bit tricky, and it is a
bit tricky to understand why writing them that way actually implements the
intend of the rules\ldots
\bparag Don't forget to update the loan Record Sheet if needed.

\subsubsection{Stability improvement}
\aparag Explain how to resolve it.
\bparag Let players compute their modifiers and decide if they want to invest
in it.
\bparag Explain that (i) a \STAB of \bonus{+2} while at peace is usually good
enough and \bonus{+3} is nice but not that nice to waste too much money on it;
(ii) it's one of those action where the immediate result can make a huge
difference, so, if in need, cough up the dough\ldots

\subsection{Turn 1 -- Peaces}
\subsubsection{Peaces}
\aparag This one is highly dependent on what happened during the turn.

\aparag Don't let \FRA make peace with \paysNaples. We want a full blown war
next turn!

\aparag Against minors, the usual choice is taking what you can now or making
a longer war (more costly, more risky) in order to rip more benefits.
\bparag When the minor is fully deployed, the increase in risk is low (no new
troops or fortress can pop up), especially in \TMED with only \fortress 2
available.
\bparag But war is never a sure thing. That last siege can fail for a full
turn and it does happen that the situation will be the same in one turn, but
the costs have increase.

\aparag \TUR probably should only sign peace by destroying \paysGeorgie.
\bparag As much as \TUR is on the rush to build its empire, it's also mostly
in control of the schedule (and its secondary forces are often enough to
react). It's likely easier to finish off \paysGeorgie now than to find time
for another war later one.
\bparag In any cases, skip the call for Crusade.

\aparag \POL and \RUS will have a harder choice between their greed and their
potential disagreements. \POL will likely not want to stop if only 1 province
can be taken, and \RUS not if only 2 provinces can be taken.
\bparag But they are a bit less in control of their schedules. They have
conflicting interests (very directly: \provinceSmolenska), other agendas
(other khanates), and other threats (\POL getting in a quagmire in the East
can pave the way of an early Turkish attack on \paysHongrie).

\aparag In both cases, it is perfectly normal to have the war go on for 2 or 3
turns. Annexing 2 or 3 provinces in one turn is not a common feat in this
game.

\subsubsection{Stability and Inflation}
\aparag Explain that war weariness (and a few other factor) decreases \STAB.
\bparag At this point, players should understand why 2 turns of war will be
common occurrence and more will be fights to death.
\bparag This might also be the good moment to mention the ``2 turns at
\bonus{-3}'' rule, since somebody is likely joking about gaming the system by
staying low all time\ldots

\aparag Explain how to compute inflation. It's again a simple but somewhat
unusual computation (because of the ``minimum as many \ducats as the
percentage value'' rule).
\bparag Mention that for most of the game, the inflation will be 25\%, not
5\%.

\subsection{Turn 1 -- Interphase}
\aparag There shouldn't be much to do here on first turn.
\bparag Count \VPs, even if you don't necessarily tally them for
tutorial. That will give players a chance to figure out what they need to do.
\bparag Mention that \VPs gain rate is highly dependent on countries, so
``winning'' after a few turns is not significant.


% Local Variables:
% fill-column: 78
% coding: utf-8-unix
% mode-require-final-newline: t
% mode: flyspell
% ispell-local-dictionary: "british"
% End: